---
title: "Pre–Post Survey Analysis"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    df_print: paged
  word_document:
    toc: true
    toc_depth: '3'
  pdf_document:
    toc: true
    toc_depth: '3'
---

```{css, echo=FALSE}
/* ── Visual polish for the HTML output ───────────────────────────────────── */
h1.title          { font-size: 2rem; font-weight: 700; }
.section-label    { color: #2c7be5; font-weight: 600; }
table             { font-size: 0.9rem; }
.sig              { color: #e63946; font-weight: bold; }
.nosig            { color: #6c757d; }
.callout          { background: #f0f4ff; border-left: 4px solid #2c7be5;
                    padding: 0.75rem 1rem; margin-bottom: 1rem;
                    border-radius: 0 4px 4px 0; }
```

---

<div class="callout">
**How to use this document**

1. Fill in the **Configuration** block in Section 1 (at minimum, verify `QUESTION_COL` matches your data).
2. Run `superset_connection.R` once to pull data from Apache Superset and cache it as CSVs.
   Alternatively, place pre-exported CSVs directly in the `data_cache/` folder.
3. Click **Knit** — the report regenerates automatically every time.

Column names are detected automatically from `pre_` / `post_` prefixes; no other code edits are required for typical datasets.
</div>

---

# Setup & Configuration {.tabset}

## Package Installation & Loading

```{r setup, message=FALSE, warning=FALSE}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 1 — SETUP
#
# This chunk:
#   1. Ensures all required R packages are installed (safe to re-run).
#   2. Loads every package into the session.
#   3. Sets global knit options (figure size, warnings, etc.).
#   4. Defines the colour palette and significance threshold used throughout.
#
# Packages used and why:
#   dplyr / tidyr  — data manipulation and reshaping
#   ggplot2        — all visualisations
#   scales         — axis formatting helpers (percent labels, etc.)
#   knitr / kableExtra — formatted tables inside the HTML/Word output
#   purrr          — functional helpers (map, keep, etc.)
#   stringr        — string manipulation (str_replace, str_to_title, etc.)
#   rstatix        — tidy-friendly statistical tests (not directly called
#                    but kept for optional extensions)
#   effectsize     — standardised effect-size calculations (e.g. Cohen's d)
# ══════════════════════════════════════════════════════════════════════════════

required_pkgs <- c(
  "dplyr", "tidyr", "ggplot2", "scales",
  "knitr", "kableExtra", "purrr", "stringr",
  "rstatix", "effectsize"
)

# Install any packages that are not yet present — runs only on first use
new_pkgs <- required_pkgs[!sapply(required_pkgs, requireNamespace, quietly = TRUE)]
if (length(new_pkgs) > 0) {
  message("Installing missing packages: ", paste(new_pkgs, collapse = ", "))
  install.packages(new_pkgs, repos = "https://cloud.r-project.org", quiet = TRUE)
}
invisible(lapply(required_pkgs, library, character.only = TRUE))

# ── Global knitr options ──────────────────────────────────────────────────────
knitr::opts_chunk$set(
  echo    = TRUE,    # show code in output (set FALSE to hide from readers)
  warning = FALSE,   # suppress package/function warnings in output
  message = FALSE,   # suppress info messages in output
  fig.width  = 9,
  fig.height = 5.5,
  dpi        = 150
)

# ── Colour palette ────────────────────────────────────────────────────────────
# Teal (pre) and orange (post) — distinguishable by most colour-blind viewers.
PAL <- c(pre = "#0077b6", post = "#f4a261")

# ── Significance threshold (α) ────────────────────────────────────────────────
# p-values below this level are flagged as statistically significant.
ALPHA <- 0.05
```

## Configuration

```{r config}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 2 — CONFIGURATION  ← EDIT THIS BLOCK
#
# All user-adjustable settings live here so that the rest of the document
# never needs to be touched between analysis runs.
# ══════════════════════════════════════════════════════════════════════════════

# Name of the column that identifies each survey question (the row label).
# Common values: "question", "item", "question_text", "label"
# If this value doesn't match your data, the script auto-detects the right
# column and prints a warning — update this value to silence that warning.
QUESTION_COL <- "question"

# Optional: a character vector of question IDs to EXCLUDE from all analyses.
# Example: EXCLUDE_QUESTIONS <- c("Q99", "warm_up_item")
EXCLUDE_QUESTIONS <- c()

# Optional: a separate column with prettier question labels for axis text.
# Set to NULL to use QUESTION_COL as-is.
LABEL_COL <- NULL

# Likert scale range used in the survey.
# Used in the average-score change plot subtitle. Adjust if using a 7-point scale.
LIKERT_MIN <- 1
LIKERT_MAX <- 5

# Report title — appears in the summary table header.
REPORT_TITLE <- "Pre–Post Survey Results"
```

---

# Helper Functions & Unit Tests {.tabset}

## Helper Definitions

```{r helpers}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 3 — HELPER FUNCTIONS
#
# All reusable utility functions are defined here, before they are called,
# so they can also be unit-tested in the next chunk.
# ══════════════════════════════════════════════════════════════════════════════

# ── Logging ───────────────────────────────────────────────────────────────────
# log_msg() prints a timestamped diagnostic line to the console.
# Use it any time a meaningful decision or data shape is worth recording.
log_msg <- function(..., level = "INFO") {
  ts <- format(Sys.time(), "%H:%M:%S")
  cat(sprintf("[%s] [%s] %s\n", ts, level, paste0(...)))
}

# ── Column-pair detection ─────────────────────────────────────────────────────
# detect_pairs() scans a data frame for columns whose names start with "pre_"
# and "post_" and returns those that form matched pairs (same stem after
# stripping the prefix). The function also reports unmatched one-sided columns
# so you can spot data-structure issues early.
#
# Returns a named list:
#   $pre      — full pre-column names of matched pairs (e.g. "pre_score")
#   $post     — full post-column names of matched pairs (e.g. "post_score")
#   $stems    — shared stems (e.g. "score")
#   $pre_only — stems found only in pre columns (no matching post)
#   $post_only— stems found only in post columns (no matching pre)
detect_pairs <- function(df) {
  pre_cols   <- grep("^pre_",  names(df), value = TRUE)
  post_cols  <- grep("^post_", names(df), value = TRUE)
  pre_stems  <- sub("^pre_",  "", pre_cols)
  post_stems <- sub("^post_", "", post_cols)
  matched    <- intersect(pre_stems, post_stems)
  list(
    pre       = paste0("pre_",  matched),
    post      = paste0("post_", matched),
    stems     = matched,
    pre_only  = setdiff(pre_stems,  post_stems),
    post_only = setdiff(post_stems, pre_stems)
  )
}

# ── Question-column auto-detection ────────────────────────────────────────────
# detect_q_col() tries to find the column that serves as the row identifier
# (survey question label). It first checks a list of common names, then falls
# back to the first column that doesn't have a pre_/post_ prefix.
detect_q_col <- function(df) {
  candidates <- c(
    "question", "Question", "question_text", "Question Text",
    "item", "Item", "label", "Label", "survey_question",
    "row_label", "name", "Name", "description", "metric"
  )
  found <- intersect(candidates, names(df))
  if (length(found) > 0) return(found[1])
  # Fall back to the first non-metric column
  other <- names(df)[!grepl("^(pre_|post_)", names(df))]
  if (length(other) > 0) return(other[1])
  return(names(df)[1])   # last resort
}

# ── Cramér's V ────────────────────────────────────────────────────────────────
# Cramér's V is a standardised effect size for chi-square tests.
# Formula: V = sqrt( chi2 / (n * (k - 1)) )
#   chi2 — chi-square statistic from chisq.test()
#   n    — total number of observations (sum of all cell counts)
#   k    — number of categories (columns in the contingency table)
# Interpretation benchmarks: 0.10 = small, 0.30 = medium, 0.50 = large
cramers_v <- function(chi2, n, k) {
  # Guard against degenerate cases that produce NaN
  if (n <= 0 || k <= 1) return(NA_real_)
  sqrt(chi2 / (n * (k - 1)))
}

# ── Safe numeric extraction ───────────────────────────────────────────────────
# safe_num() converts a value to numeric, returning NA rather than throwing
# an error when conversion fails. This handles cells that contain empty
# strings, "N/A", or other non-numeric text from Superset exports.
safe_num <- function(x) suppressWarnings(as.numeric(x))

# ── Significance star formatter ───────────────────────────────────────────────
# sig_flag() converts a logical (TRUE/FALSE/NA) into a readable flag string.
sig_flag <- function(x) {
  dplyr::case_when(
    is.na(x)  ~ "—",
    x == TRUE ~ "✓ Yes",
    TRUE      ~ "No"
  )
}
```

## Unit Tests

```{r unit-tests}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 4 — UNIT TESTS
#
# These tests run automatically every time the document is knitted.
# They verify that helper functions behave correctly with known inputs.
# A failure here means a helper function was accidentally broken —
# check any recent edits to the helpers chunk above.
#
# Each test uses stopifnot(), which halts execution with a clear message
# if the assertion is FALSE. The cat() lines provide progress visibility.
# ══════════════════════════════════════════════════════════════════════════════

cat("Running unit tests for helper functions...\n\n")

# ── Test: detect_pairs() ─────────────────────────────────────────────────────
cat("  [TEST] detect_pairs()\n")
test_df_pairs <- data.frame(
  question   = c("Q1", "Q2"),
  pre_score  = c(3.2, 4.1),
  post_score = c(4.0, 4.5),
  pre_count  = c(50L, 48L),
  post_count = c(52L, 49L),
  pre_delta  = c(0.8, 0.4)   # has pre_ prefix but no post_delta → lands in pre_only
  # Note: detect_pairs only sees columns starting with pre_/post_, so a bare
  # column named "delta" (no prefix) is invisible to it — hence pre_delta here.
)
pairs_test <- detect_pairs(test_df_pairs)
stopifnot("detect_pairs finds 2 matched stems" =
            length(pairs_test$stems) == 2)
stopifnot("detect_pairs identifies 'score' stem" =
            "score" %in% pairs_test$stems)
stopifnot("detect_pairs identifies 'count' stem" =
            "count" %in% pairs_test$stems)
stopifnot("detect_pairs flags 'delta' as pre_only" =
            "delta" %in% pairs_test$pre_only)
stopifnot("detect_pairs pre_only and post_only are disjoint" =
            length(intersect(pairs_test$pre_only, pairs_test$post_only)) == 0)
cat("    PASSED\n")

# ── Test: detect_q_col() ─────────────────────────────────────────────────────
cat("  [TEST] detect_q_col()\n")
stopifnot("detects 'question' by name" =
            detect_q_col(data.frame(question="a", pre_x=1)) == "question")
stopifnot("detects 'item' when 'question' absent" =
            detect_q_col(data.frame(item="a", pre_x=1)) == "item")
stopifnot("falls back to first non-pre/post col" =
            detect_q_col(data.frame(pre_x=1, post_x=2, mykey="a")) == "mykey")
cat("    PASSED\n")

# ── Test: cramers_v() ────────────────────────────────────────────────────────
cat("  [TEST] cramers_v()\n")
# Known result: sqrt(10 / (100 * 2)) ≈ 0.2236
expected_v <- sqrt(10 / (100 * 2))
stopifnot("cramers_v produces correct numeric result" =
            abs(cramers_v(10, 100, 3) - expected_v) < 1e-10)
stopifnot("cramers_v returns NA for degenerate n=0" =
            is.na(cramers_v(10, 0, 3)))
stopifnot("cramers_v returns NA for degenerate k=1" =
            is.na(cramers_v(10, 100, 1)))
cat("    PASSED\n")

# ── Test: safe_num() ─────────────────────────────────────────────────────────
cat("  [TEST] safe_num()\n")
stopifnot("safe_num converts numeric string" = safe_num("3.14") == 3.14)
stopifnot("safe_num returns NA for non-numeric" = is.na(safe_num("N/A")))
stopifnot("safe_num handles empty string" = is.na(safe_num("")))
stopifnot("safe_num passes through numeric" = safe_num(42) == 42)
cat("    PASSED\n")

# ── Test: sig_flag() ─────────────────────────────────────────────────────────
cat("  [TEST] sig_flag()\n")
stopifnot("sig_flag TRUE → '✓ Yes'" = sig_flag(TRUE)  == "✓ Yes")
stopifnot("sig_flag FALSE → 'No'"   = sig_flag(FALSE) == "No")
stopifnot("sig_flag NA → '—'"       = sig_flag(NA)    == "—")
cat("    PASSED\n")

cat("\nAll unit tests passed.\n")
```

---

# Data Import & Validation

```{r import-data}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 5 — DATA IMPORT
#
# Data is read from CSV files cached by superset_connection.R.
# To refresh from Superset, uncomment the source() line below and comment
# out the read.csv() block — or just re-run superset_connection.R first.
#
# check.names = FALSE preserves column names exactly as exported from
# Superset (spaces, special characters, etc. are kept intact).
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Starting data import from data_cache/")

# ── Option A (live pull) ──────────────────────────────────────────────────────
# Uncomment to pull fresh data from Superset before loading:
# source("superset_connection.R")

# ── Option B (cached CSVs — default) ─────────────────────────────────────────
cache_path  <- "data_cache"
cache_files <- list(
  slider      = file.path(cache_path, "slider.csv"),
  knowledge   = file.path(cache_path, "knowledge.csv"),
  likert      = file.path(cache_path, "likert.csv"),
  multiselect = file.path(cache_path, "multiselect.csv")
)

# Verify all cache files exist before attempting to read them
missing_files <- names(cache_files)[!file.exists(unlist(cache_files))]
if (length(missing_files) > 0) {
  stop(
    "Missing cache file(s): ", paste(missing_files, collapse = ", "), "\n",
    "Run superset_connection.R first to generate the CSVs, ",
    "or place them manually in the '", cache_path, "/' folder."
  )
}

read_cache <- function(path) {
  read.csv(path, stringsAsFactors = FALSE, check.names = FALSE)
}

df_slider      <- read_cache(cache_files$slider)
df_knowledge   <- read_cache(cache_files$knowledge)
df_likert      <- read_cache(cache_files$likert)
df_multiselect <- read_cache(cache_files$multiselect)

log_msg("Cache files loaded successfully")
```

```{r validate-data}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 6 — DATA VALIDATION
#
# Before running any analysis, we verify that the data meets basic
# structural requirements. Problems are reported as warnings (non-fatal)
# so that the rest of the report still renders — making it easier to
# diagnose issues incrementally.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Running data validation checks")

# ── Helper: validate a single data frame ─────────────────────────────────────
validate_df <- function(df, name) {
  issues <- character(0)

  if (nrow(df) == 0)
    issues <- c(issues, "has 0 rows — data may not have loaded correctly")

  if (ncol(df) == 0)
    issues <- c(issues, "has 0 columns — check the CSV file")

  # Flag columns that are entirely NA (useless for analysis)
  all_na_cols <- names(df)[sapply(df, function(col) all(is.na(col)))]
  if (length(all_na_cols) > 0)
    issues <- c(issues, paste("all-NA columns:", paste(all_na_cols, collapse=", ")))

  # Flag numeric-looking columns stored as character (common Superset export issue)
  # vapply() is used instead of sapply() because it guarantees a logical vector
  # as output — sapply() can silently return a list when columns loaded with
  # check.names=FALSE have unusual types, causing "invalid subscript type 'list'".
  # tryCatch() inside ensures any single problematic column returns FALSE safely.
  pre_post_cols <- names(df)[grepl("^(pre_|post_)", names(df))]
  char_num_cols <- if (length(pre_post_cols) == 0) character(0) else {
    pre_post_cols[
      vapply(pre_post_cols, function(col) {
        tryCatch({
          vals <- df[[col]]
          is.character(vals) && !all(is.na(suppressWarnings(as.numeric(vals))))
        }, error = function(e) FALSE)
      }, logical(1))
    ]
  }
  if (length(char_num_cols) > 0)
    issues <- c(issues, paste("numeric values stored as character:",
                              paste(char_num_cols, collapse=", "),
                              "— conversion will be applied automatically"))

  if (length(issues) > 0) {
    for (iss in issues)
      log_msg(name, ": ", iss, level = "WARN")
  } else {
    log_msg(name, ": OK (", nrow(df), " rows × ", ncol(df), " cols)")
  }

  invisible(df)
}

validate_df(df_slider,      "df_slider")
validate_df(df_knowledge,   "df_knowledge")
validate_df(df_likert,      "df_likert")
validate_df(df_multiselect, "df_multiselect")

# ── Auto-detect QUESTION_COL ─────────────────────────────────────────────────
# If the configured QUESTION_COL doesn't exist in the knowledge data frame
# (the primary reference), we scan for a sensible alternative.
if (!QUESTION_COL %in% names(df_knowledge)) {
  detected <- detect_q_col(df_knowledge)
  log_msg(
    "QUESTION_COL '", QUESTION_COL, "' not found — auto-detected: '",
    detected, "'  |  Update QUESTION_COL in Section 1 to silence this.",
    level = "WARN"
  )
  QUESTION_COL <- detected
}
log_msg("QUESTION_COL set to: '", QUESTION_COL, "'")

# Warn if any dataset is missing QUESTION_COL (non-fatal — row index used instead)
for (nm in c("df_slider", "df_knowledge", "df_likert", "df_multiselect")) {
  df <- get(nm)
  if (!QUESTION_COL %in% names(df))
    log_msg(nm, " does not contain QUESTION_COL '", QUESTION_COL,
            "' — row index will be used as label", level = "WARN")
}

# ── Apply exclusions ──────────────────────────────────────────────────────────
# Remove any questions listed in EXCLUDE_QUESTIONS from all data frames.
# This is useful for filtering out pilot questions, demographic items, etc.
exclude_rows <- function(df) {
  if (length(EXCLUDE_QUESTIONS) > 0 && QUESTION_COL %in% names(df))
    df[!df[[QUESTION_COL]] %in% EXCLUDE_QUESTIONS, , drop = FALSE]
  else
    df
}

df_slider      <- exclude_rows(df_slider)
df_knowledge   <- exclude_rows(df_knowledge)
df_likert      <- exclude_rows(df_likert)
df_multiselect <- exclude_rows(df_multiselect)

if (length(EXCLUDE_QUESTIONS) > 0)
  log_msg("Excluded ", length(EXCLUDE_QUESTIONS), " question(s): ",
          paste(EXCLUDE_QUESTIONS, collapse = ", "))

# ── Dimension summary ─────────────────────────────────────────────────────────
cat("\nDataset dimensions after validation and exclusions:\n")
cat(sprintf("  %-16s : %d rows × %d cols\n", "df_slider",      nrow(df_slider),      ncol(df_slider)))
cat(sprintf("  %-16s : %d rows × %d cols\n", "df_knowledge",   nrow(df_knowledge),   ncol(df_knowledge)))
cat(sprintf("  %-16s : %d rows × %d cols\n", "df_likert",      nrow(df_likert),      ncol(df_likert)))
cat(sprintf("  %-16s : %d rows × %d cols\n", "df_multiselect", nrow(df_multiselect), ncol(df_multiselect)))
```

## Raw Data Preview

```{r data-preview}
# Quick look at the first few rows of each dataset.
# This is useful for confirming column names match what the analysis expects.
cat("── Perception Slider ──\n");  print(head(df_slider))
cat("\n── Knowledge ──\n");        print(head(df_knowledge))
cat("\n── Likert ──\n");           print(head(df_likert))
cat("\n── Multi-Select ──\n");     print(head(df_multiselect))
```

---

# Column Auto-Detection

```{r detect-pairs}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 7 — COLUMN AUTO-DETECTION
#
# We use the detect_pairs() helper (defined in Section 3) to identify every
# pre_/post_ column pair in each dataset. The "stem" is the shared part of
# the column name after stripping the prefix.
#
# Example: "pre_avg_score" and "post_avg_score" → stem = "avg_score"
#
# Columns without a matching counterpart are flagged as pre-only or post-only
# so you can quickly spot data structure issues.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Detecting pre_/post_ column pairs")

pairs_s <- detect_pairs(df_slider)
pairs_k <- detect_pairs(df_knowledge)
pairs_l <- detect_pairs(df_likert)
pairs_m <- detect_pairs(df_multiselect)

dataset_names <- c(Slider = "pairs_s", Knowledge = "pairs_k",
                   Likert = "pairs_l", `Multi-Select` = "pairs_m")

for (label in names(dataset_names)) {
  p <- get(dataset_names[[label]])
  cat(label, "— matched pairs (", length(p$stems), "):",
      if (length(p$stems) > 0) paste(p$stems, collapse=", ") else "(none)", "\n")
  if (length(p$pre_only)  > 0)
    cat("    pre-only  :", paste(p$pre_only,  collapse=", "), "\n")
  if (length(p$post_only) > 0)
    cat("    post-only :", paste(p$post_only, collapse=", "), "\n")
}
```

```{r classify-columns}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 8 — COLUMN TYPE CLASSIFICATION
#
# Within each dataset, stems are further classified by their role
# (score/average, count, percentage, delta) using regex pattern matching.
# This drives which statistical tests and visualisations are applied.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Classifying column types within each dataset")

# ── Perception Slider ─────────────────────────────────────────────────────────
# Score columns contain average/mean values (continuous, used for comparisons).
# Count columns hold numeric tallies (used if an n-based t-test is possible).
s_score_stems <- pairs_s$stems[grepl("avg|mean|score|value",
                                     pairs_s$stems, ignore.case = TRUE)]
s_count_stems <- pairs_s$stems[!grepl("avg|mean|score|value|pct|percent|delta",
                                      pairs_s$stems, ignore.case = TRUE)]

# ── Knowledge ─────────────────────────────────────────────────────────────────
# count_stems: raw correct/incorrect/total counts  → used in prop.test
# pct_stems:   percentages                          → reported descriptively
# delta_cols:  pre-computed change columns          → reported descriptively
k_count_stems <- pairs_k$stems[!grepl("pct|percent|delta|avg|mean|score",
                                      pairs_k$stems, ignore.case = TRUE)]
k_pct_stems   <- pairs_k$stems[ grepl("pct|percent",
                                      pairs_k$stems, ignore.case = TRUE)]
k_delta_cols  <- grep("delta", names(df_knowledge), value = TRUE, ignore.case = TRUE)

# ── Likert ────────────────────────────────────────────────────────────────────
# Category stems (strongly_disagree … strongly_agree) form the contingency
# table for the chi-square test.
# Score stems hold the average Likert score per question.
likert_cat_order <- c("strongly_disagree", "disagree", "neutral",
                      "agree", "strongly_agree")
l_cat_stems  <- intersect(likert_cat_order, pairs_l$stems)
l_score_stem <- pairs_l$stems[grepl("avg|mean|score", pairs_l$stems,
                                    ignore.case = TRUE)]

# ── Multi-Select ──────────────────────────────────────────────────────────────
# Option stems are the individual response options (each treated as a binary).
# The total stem holds the total respondent count (denominator for percentages).
m_option_stems <- pairs_m$stems[!grepl("pct|percent|delta|avg|mean|total",
                                       pairs_m$stems, ignore.case = TRUE)]
m_total_stem   <- pairs_m$stems[grepl("total", pairs_m$stems, ignore.case = TRUE)]
m_total_stem   <- if (length(m_total_stem) > 0) m_total_stem[1] else NA_character_

# ── Print classification summary ──────────────────────────────────────────────
cat("\nSlider:\n")
cat("  score stems  :", paste0(s_score_stems, collapse=", "), "\n")
cat("  count stems  :", paste0(s_count_stems, collapse=", "), "\n")
cat("\nKnowledge:\n")
cat("  count stems  :", paste(k_count_stems, collapse=", "), "\n")
cat("  pct stems    :", paste(k_pct_stems,   collapse=", "), "\n")
cat("  delta cols   :", paste(k_delta_cols,  collapse=", "), "\n")
cat("\nLikert:\n")
cat("  category stems:", paste(l_cat_stems,  collapse=", "), "\n")
cat("  score stems   :", paste(l_score_stem, collapse=", "), "\n")
cat("\nMulti-Select:\n")
cat("  option stems :", paste(m_option_stems, collapse=", "), "\n")
cat("  total stem   :", m_total_stem, "\n")

log_msg("Column classification complete")
```

---

# Perception Slider Analysis

> **Statistical method:** Descriptive comparison of pre vs. post mean scores
> per question. Because only aggregate means are available (not individual
> responses), a formal paired t-test cannot be applied. If your data also
> contains standard deviation and sample-size columns, a two-sample Welch
> t-test is run from the summary statistics. Otherwise, results are reported
> descriptively with the raw change noted.

## Statistical Tests

```{r slider-stats}
# ══════════════════════════════════════════════════════════════════════════════
# Perception Slider: mean-score comparison
#
# For each question row we compare the pre and post mean scores.
# If SD and N columns are present, a two-sample Welch t-test is computed
# from the group summary statistics (mean, SD, n) rather than raw data.
#
# The Welch t-test does not assume equal variances between groups, making
# it more robust than a standard Student's t-test for aggregate data.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Running Perception Slider analysis")

# Identify the primary score column (first detected score stem)
pre_s_col  <- if (length(s_score_stems) > 0) paste0("pre_",  s_score_stems[1]) else NULL
post_s_col <- if (length(s_score_stems) > 0) paste0("post_", s_score_stems[1]) else NULL

# Check for optional SD and N columns (enable t-test if found)
pre_n_col   <- grep("^pre_(n|count|total|respondents)$",
                    names(df_slider), value = TRUE, ignore.case = TRUE)[1]
post_n_col  <- grep("^post_(n|count|total|respondents)$",
                    names(df_slider), value = TRUE, ignore.case = TRUE)[1]
pre_sd_col  <- grep("^pre_(sd|std|stdev)",
                    names(df_slider), value = TRUE, ignore.case = TRUE)[1]
post_sd_col <- grep("^post_(sd|std|stdev)",
                    names(df_slider), value = TRUE, ignore.case = TRUE)[1]

has_ttest_inputs <- !is.na(pre_n_col) && !is.na(post_n_col) &&
                    !is.na(pre_sd_col) && !is.na(post_sd_col)

log_msg("Slider — score cols: ", pre_s_col, " / ", post_s_col,
        " | t-test inputs available: ", has_ttest_inputs)

# ── Per-question test function ────────────────────────────────────────────────
run_slider_test <- function(row) {
  # Safe label extraction (falls back to row index if QUESTION_COL is absent)
  q_label   <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
                 as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))
  pre_mean  <- safe_num(row[[pre_s_col]])
  post_mean <- safe_num(row[[post_s_col]])
  diff      <- round(post_mean - pre_mean, 3)

  if (has_ttest_inputs) {
    # Two-sample Welch t-test from summary statistics
    n1  <- safe_num(row[[pre_n_col]]);  n2  <- safe_num(row[[post_n_col]])
    sd1 <- safe_num(row[[pre_sd_col]]); sd2 <- safe_num(row[[post_sd_col]])
    se   <- sqrt(sd1^2 / n1 + sd2^2 / n2)
    t    <- diff / se
    df_t <- (sd1^2/n1 + sd2^2/n2)^2 /
            ((sd1^2/n1)^2/(n1-1) + (sd2^2/n2)^2/(n2-1))
    p    <- 2 * pt(-abs(t), df_t)
    # Cohen's d using pooled SD
    sp   <- sqrt(((n1-1)*sd1^2 + (n2-1)*sd2^2) / (n1+n2-2))
    d    <- round(diff / sp, 3)
    data.frame(question=q_label,
               pre_mean=round(pre_mean,2), post_mean=round(post_mean,2),
               diff=diff, p_value=round(p,4), cohens_d=d,
               significant=p < ALPHA, method="Welch t-test",
               stringsAsFactors=FALSE)
  } else {
    # Descriptive only — Cohen's d cannot be computed without SD
    data.frame(question=q_label,
               pre_mean=round(pre_mean,2), post_mean=round(post_mean,2),
               diff=diff, p_value=NA_real_, cohens_d=NA_real_,
               significant=NA, method="Descriptive (no SD available)",
               stringsAsFactors=FALSE)
  }
}

# ── Run across all rows ───────────────────────────────────────────────────────
if (is.null(pre_s_col)) {
  log_msg("No score columns found in df_slider — skipping slider analysis", level="WARN")
  results_slider <- data.frame()
} else {
  results_slider <- bind_rows(
    lapply(seq_len(nrow(df_slider)), function(i) run_slider_test(df_slider[i, ]))
  )
  n_sig <- sum(results_slider$significant == TRUE, na.rm = TRUE)
  log_msg("Slider: ", nrow(results_slider), " questions | ", n_sig, " significant")

  # ── Results table ───────────────────────────────────────────────────────────
  results_slider %>%
    mutate(
      diff     = paste0(ifelse(diff >= 0, "+", ""), diff),
      sig_flag = sig_flag(significant)
    ) %>%
    select(question, pre_mean, post_mean, diff, p_value, cohens_d, sig_flag, method) %>%
    kable(
      col.names = c("Question", "Pre Mean", "Post Mean", "Change",
                    "p-value", "Cohen's d", "Significant?", "Method"),
      caption   = "Perception Slider: Pre vs. Post Mean Score Comparison"
    ) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE) %>%
    row_spec(which(results_slider$significant == TRUE),
             bold = TRUE, color = "#e63946")
}
```

## Visualisation — Score Change

```{r slider-plot, fig.height=max(4, nrow(df_slider) * 0.55 + 1.5)}
# Grouped horizontal bar chart comparing pre vs. post mean score per question.
# Questions with a statistically significant change are starred (*).
if (nrow(results_slider) > 0) {
  # Reshape to long format for ggplot (one row per question × time point)
  plot_s <- results_slider %>%
    select(question, pre_mean, post_mean) %>%
    pivot_longer(c(pre_mean, post_mean), names_to = "time", values_to = "score") %>%
    mutate(
      time = recode(time, pre_mean = "Pre", post_mean = "Post"),
      time = factor(time, levels = c("Pre", "Post"))
    )

  # Add a "*" to labels for statistically significant questions
  sig_lbl_s <- results_slider %>%
    mutate(label = ifelse(!is.na(significant) & significant,
                          paste0(question, " *"), question)) %>%
    select(question, label)
  plot_s <- left_join(plot_s, sig_lbl_s, by = "question")

  ggplot(plot_s, aes(x = score, y = reorder(label, score), fill = time)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6, colour = "white") +
    geom_text(aes(label = round(score, 2)),
              position = position_dodge(width = 0.7), hjust = -0.15, size = 3.2) +
    scale_fill_manual(values = PAL, name = "Time point") +
    scale_x_continuous(expand = expansion(mult = c(0, 0.2))) +
    labs(
      title    = "Perception Slider: Mean Score Pre vs. Post",
      subtitle = "* statistically significant change (p < 0.05)",
      x = "Mean Score", y = NULL
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top", panel.grid.major.y = element_blank())
}
```

---

# Knowledge Questions Analysis

> **Statistical method:** Two-proportion z-test (`prop.test`) comparing
> the proportion of correct responses at pre vs. post for each question.
> This test is appropriate when we have aggregate counts (correct, total)
> rather than individual-level binary responses.
> Effect size is reported as the absolute percentage-point change.

## Statistical Tests

```{r knowledge-stats}
# ══════════════════════════════════════════════════════════════════════════════
# Knowledge: two-proportion z-test
#
# For each question we compute:
#   Pre proportion  = pre_correct  / pre_total
#   Post proportion = post_correct / post_total
#
# prop.test() is equivalent to a chi-square test on a 2×2 table and produces
# a p-value for H₀: p_pre = p_post (proportions are equal).
#
# Note: if your data only has pre_total (no post_total), the script reuses
# pre_total as the denominator for both time points — valid when the same
# cohort answered both waves.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Running Knowledge analysis")

# ── Identify correct / total column stems ─────────────────────────────────────
correct_stem <- k_count_stems[grepl("correct",  k_count_stems, ignore.case = TRUE)][1]
total_stem   <- k_count_stems[grepl("total",    k_count_stems, ignore.case = TRUE)][1]

# ── Fallback: look for total in pre-only columns ──────────────────────────────
# Handles the common case where "pre_total" exists but "post_total" was not
# exported, because the same cohort answered both survey waves.
if (is.na(total_stem) && length(pairs_k$pre_only) > 0) {
  pre_only_total <- pairs_k$pre_only[grepl("total", pairs_k$pre_only, ignore.case = TRUE)]
  if (length(pre_only_total) > 0) {
    total_stem <- pre_only_total[1]
    log_msg("No post_total found — reusing pre_total for both denominators", level = "WARN")
  }
}

# Halt if we still can't find the essential columns
if (is.na(correct_stem) || is.na(total_stem)) {
  stop(
    "Could not auto-detect 'correct' and 'total' count columns in df_knowledge.\n",
    "Available stems (matched pairs): ", paste(k_count_stems, collapse=", "), "\n",
    "Pre-only stems: ", paste(pairs_k$pre_only, collapse=", "), "\n",
    "Ensure your knowledge data has columns named like: ",
    "pre_correct, post_correct, pre_total (and optionally post_total)."
  )
}

# Build full column names from stems
pre_correct_col  <- paste0("pre_",  correct_stem)
post_correct_col <- paste0("post_", correct_stem)
pre_total_col    <- paste0("pre_",  total_stem)
# Use post_total if it exists, otherwise fall back to pre_total
post_total_col   <- if (paste0("post_", total_stem) %in% names(df_knowledge))
                      paste0("post_", total_stem) else pre_total_col

log_msg("Knowledge — correct: ", pre_correct_col, "/", post_correct_col,
        " | total: ", pre_total_col, "/", post_total_col)

# ── Per-question prop.test ────────────────────────────────────────────────────
run_prop_test <- function(row) {
  q_label <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
               as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))

  x1 <- safe_num(row[[pre_correct_col]])
  n1 <- safe_num(row[[pre_total_col]])
  x2 <- safe_num(row[[post_correct_col]])
  n2 <- safe_num(row[[post_total_col]])

  # Skip rows where required values are missing or denominator is zero
  if (any(is.na(c(x1, n1, x2, n2))) || n1 == 0 || n2 == 0)
    return(data.frame(
      question=q_label, pre_pct=NA_real_, post_pct=NA_real_,
      pct_point_diff=NA_real_, p_value=NA_real_,
      significant=NA, note="Missing data", stringsAsFactors=FALSE
    ))

  test <- tryCatch(
    prop.test(c(x1, x2), c(n1, n2), correct = FALSE),
    error = function(e) NULL
  )

  data.frame(
    question       = q_label,
    pre_pct        = round(x1 / n1 * 100, 1),
    post_pct       = round(x2 / n2 * 100, 1),
    pct_point_diff = round((x2/n2 - x1/n1) * 100, 1),   # positive = improvement
    p_value        = if (!is.null(test)) round(test$p.value, 4) else NA_real_,
    significant    = if (!is.null(test)) test$p.value < ALPHA else NA,
    note           = "",
    stringsAsFactors = FALSE
  )
}

results_knowledge <- bind_rows(
  lapply(seq_len(nrow(df_knowledge)), function(i) run_prop_test(df_knowledge[i, ]))
)

n_sig_k <- sum(results_knowledge$significant == TRUE, na.rm = TRUE)
log_msg("Knowledge: ", nrow(results_knowledge), " questions | ", n_sig_k, " significant")

# ── Results table ─────────────────────────────────────────────────────────────
results_knowledge %>%
  mutate(
    pre_pct_fmt  = paste0(pre_pct,  "%"),
    post_pct_fmt = paste0(post_pct, "%"),
    change_fmt   = paste0(ifelse(pct_point_diff >= 0, "+", ""),
                          pct_point_diff, "pp"),
    sig          = sig_flag(significant)
  ) %>%
  select(question, pre_pct_fmt, post_pct_fmt, change_fmt, p_value, sig) %>%
  kable(
    col.names = c("Question", "Pre %", "Post %", "Change", "p-value", "Significant?"),
    caption   = "Two-proportion z-test: pre vs. post % correct",
    align     = "lccccc"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  row_spec(which(results_knowledge$significant == TRUE),
           bold = TRUE, color = "#e63946")
```

## Visualisation — % Correct Pre vs. Post

```{r knowledge-plot-bar, fig.height=max(4, nrow(results_knowledge) * 0.55 + 1.5)}
# Grouped horizontal bar chart.
# Each question has two bars side-by-side (pre = teal, post = orange).
# Stars mark questions with a statistically significant change.

plot_k <- results_knowledge %>%
  select(question, pre_pct, post_pct, significant) %>%
  # Create a display label: add " *" for significant questions
  mutate(label = ifelse(significant == TRUE, paste0(question, " *"), question)) %>%
  pivot_longer(c(pre_pct, post_pct), names_to = "time", values_to = "pct_correct") %>%
  mutate(
    time = recode(time, pre_pct = "Pre", post_pct = "Post"),
    time = factor(time, levels = c("Pre", "Post"))
  )

ggplot(plot_k, aes(x = pct_correct, y = reorder(label, pct_correct), fill = time)) +
  geom_col(position = position_dodge(width = 0.7), width = 0.6, colour = "white") +
  geom_text(aes(label = paste0(round(pct_correct, 1), "%")),
            position = position_dodge(width = 0.7), hjust = -0.15, size = 3.2) +
  scale_fill_manual(values = PAL, name = "Time point") +
  scale_x_continuous(labels = percent_format(scale = 1),
                     expand = expansion(mult = c(0, 0.18))) +
  labs(
    title    = "Knowledge Questions: % Correct Responses",
    subtitle = "* indicates statistically significant change (p < 0.05)",
    x = "% Correct", y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 10))
```

## Visualisation — Percentage-Point Change (Dot Plot)

```{r knowledge-plot-dot, fig.height=max(4, nrow(results_knowledge) * 0.55 + 1.5)}
# Lollipop / dot plot showing the magnitude of change (Post − Pre) per question.
# Questions to the right of zero improved; left of zero declined.
# Red dots indicate statistically significant differences.

ggplot(results_knowledge,
       aes(x = pct_point_diff,
           y = reorder(question, pct_point_diff),
           colour = significant)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_segment(aes(x = 0, xend = pct_point_diff,
                   y = question, yend = question),
               colour = "grey80", linewidth = 0.8) +
  geom_point(size = 4) +
  geom_text(
    aes(
      label = paste0(ifelse(pct_point_diff >= 0, "+", ""), pct_point_diff, "pp"),
      hjust = ifelse(pct_point_diff >= 0, -0.3, 1.3)
    ),
    size = 3.2
  ) +
  scale_colour_manual(
    values  = c("TRUE" = "#e63946", "FALSE" = "#6c757d"),
    labels  = c("TRUE" = "Significant (p<.05)", "FALSE" = "Not significant"),
    na.value = "grey80",
    name    = NULL
  ) +
  scale_x_continuous(
    labels = function(x) paste0(ifelse(x >= 0, "+", ""), x, "pp"),
    expand = expansion(mult = 0.25)
  ) +
  labs(
    title    = "Knowledge Questions: Percentage-Point Change (Post − Pre)",
    subtitle = "Positive = improvement; red = statistically significant",
    x = "Percentage-point change", y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top", panel.grid.major.y = element_blank())
```

---

# Likert Questions Analysis

> **Statistical method:** Chi-square test comparing the *frequency distribution*
> of responses across Likert categories (e.g. Strongly Disagree … Strongly Agree)
> at pre vs. post for each question. We use the distribution rather than just the
> mean because Likert data is ordinal — the distances between categories are not
> guaranteed to be equal. Effect size is reported as Cramér's V. Average-score
> comparisons are descriptive only (no individual-level paired data available).

## Statistical Tests

```{r likert-stats}
# ══════════════════════════════════════════════════════════════════════════════
# Likert: chi-square on response distribution
#
# For each question we build a 2 × k contingency table:
#       | SD | D | N | A | SA |
#   Pre |    |   |   |   |    |
#  Post |    |   |   |   |    |
#
# chisq.test() tests whether the response distribution shifted between waves.
# When any cell count is < 5 (violating the chi-square approximation),
# simulate.p.value=TRUE uses a Monte Carlo approximation instead.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Running Likert analysis")

# Build category column name vectors from detected stems
pre_cat_cols  <- paste0("pre_",  l_cat_stems)
post_cat_cols <- paste0("post_", l_cat_stems)
pre_score_col  <- if (length(l_score_stem) > 0) paste0("pre_",  l_score_stem[1]) else NULL
post_score_col <- if (length(l_score_stem) > 0) paste0("post_", l_score_stem[1]) else NULL

log_msg("Likert — category cols: ", length(l_cat_stems),
        " | score col: ", if(!is.null(pre_score_col)) pre_score_col else "none")

# ── Per-question chi-square ───────────────────────────────────────────────────
run_chisq <- function(row) {
  q_label <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
               as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))

  pre_counts  <- safe_num(unlist(row[pre_cat_cols]))
  post_counts <- safe_num(unlist(row[post_cat_cols]))

  # Drop categories where either pre or post is NA
  # (some questions may not use all Likert levels)
  valid       <- !is.na(pre_counts) & !is.na(post_counts)
  pre_counts  <- pre_counts[valid]
  post_counts <- post_counts[valid]

  if (length(pre_counts) < 2 || sum(pre_counts) == 0 || sum(post_counts) == 0)
    return(data.frame(
      question=q_label, pre_avg=NA_real_, post_avg=NA_real_, avg_diff=NA_real_,
      chi_sq=NA_real_, df=NA_integer_, p_value=NA_real_, cramers_v=NA_real_,
      significant=NA, note="Insufficient data", stringsAsFactors=FALSE
    ))

  mat  <- rbind(pre_counts, post_counts)
  # Use Monte Carlo p-value when any cell < 5 (expected frequency assumption)
  test <- tryCatch(
    chisq.test(mat, simulate.p.value = any(mat < 5)),
    error = function(e) NULL
  )

  n_pre  <- sum(pre_counts)
  n_post <- sum(post_counts)
  k      <- length(pre_counts)

  pre_avg  <- if (!is.null(pre_score_col))
                safe_num(row[[pre_score_col]]) else NA_real_
  post_avg <- if (!is.null(post_score_col))
                safe_num(row[[post_score_col]]) else NA_real_

  data.frame(
    question    = q_label,
    pre_avg     = round(pre_avg,  2),
    post_avg    = round(post_avg, 2),
    avg_diff    = round(post_avg - pre_avg, 2),
    chi_sq      = if (!is.null(test)) round(test$statistic, 3) else NA_real_,
    df          = if (!is.null(test)) as.integer(test$parameter) else NA_integer_,
    p_value     = if (!is.null(test)) round(test$p.value, 4)  else NA_real_,
    cramers_v   = if (!is.null(test))
                    round(cramers_v(test$statistic, n_pre + n_post, k), 3)
                  else NA_real_,
    significant = if (!is.null(test)) test$p.value < ALPHA else NA,
    note        = if (!is.null(test) && any(mat < 5))
                    "Simulated p-value (low cell counts)" else "",
    stringsAsFactors = FALSE
  )
}

results_likert <- bind_rows(
  lapply(seq_len(nrow(df_likert)), function(i) run_chisq(df_likert[i, ]))
)

n_sig_l <- sum(results_likert$significant == TRUE, na.rm = TRUE)
log_msg("Likert: ", nrow(results_likert), " questions | ", n_sig_l, " significant")

# ── Results table ─────────────────────────────────────────────────────────────
results_likert %>%
  mutate(
    avg_diff_fmt = paste0(ifelse(!is.na(avg_diff) & avg_diff >= 0, "+", ""), avg_diff),
    sig          = sig_flag(significant)
  ) %>%
  select(question, pre_avg, post_avg, avg_diff_fmt,
         chi_sq, df, p_value, cramers_v, sig) %>%
  kable(
    col.names = c("Question", "Pre Avg", "Post Avg", "Δ Avg",
                  "χ²", "df", "p-value", "Cramér's V", "Significant?"),
    caption   = paste0(
      "Chi-square test: distribution of Likert responses pre vs. post\n",
      "(Cramér's V interpretation: 0.10 = small, 0.30 = medium, 0.50 = large effect)"
    ),
    align = "lcccccccc"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"),
                full_width = FALSE) %>%
  row_spec(which(results_likert$significant == TRUE),
           bold = TRUE, color = "#e63946") %>%
  {
    # Add footnote only when at least one note exists (avoids empty footnote block)
    notes <- results_likert$note[nchar(results_likert$note) > 0]
    if (length(notes) > 0)
      footnote(., general = notes, general_title = "Notes: ")
    else .
  }
```

## Visualisation — Diverging Stacked Bar Chart

```{r likert-plot-diverging, fig.height=max(5, nrow(df_likert) * 1.2 + 2)}
# ══════════════════════════════════════════════════════════════════════════════
# Diverging stacked bar chart
#
# This is the standard visualisation for Likert data. Responses are centred
# at zero (the "neutral" point). Disagree/Strongly Disagree extend left;
# Agree/Strongly Agree extend right. Neutral is split half-left/half-right
# so bars are comparable between questions.
#
# Each question has two rows: Pre and Post (separated by the question label).
# ══════════════════════════════════════════════════════════════════════════════

if (length(l_cat_stems) == 0) {
  message("No Likert category columns detected — skipping diverging chart.")
} else {

# ── Build long format ─────────────────────────────────────────────────────────
build_diverging <- function(df, time_label, cat_cols, cat_stems) {
  df %>%
    select(all_of(c(QUESTION_COL, cat_cols))) %>%
    pivot_longer(cols = all_of(cat_cols), names_to = "category", values_to = "count") %>%
    mutate(
      # Strip the "pre_" / "post_" prefix to get the bare category name
      category = str_remove(category, paste0("^", time_label, "_")),
      time     = time_label,
      count    = safe_num(count)
    )
}

long_pre  <- build_diverging(df_likert, "pre",  pre_cat_cols,  l_cat_stems)
long_post <- build_diverging(df_likert, "post", post_cat_cols, l_cat_stems)

long_all <- bind_rows(long_pre, long_post) %>%
  mutate(
    time     = factor(time, levels = c("pre", "post"), labels = c("Pre", "Post")),
    category = factor(category,
                      levels = rev(likert_cat_order[likert_cat_order %in% l_cat_stems]))
  )

# ── Convert to row percentages ─────────────────────────────────────────────────
# We show % of respondents rather than raw counts so that questions with
# different numbers of respondents are directly comparable.
long_all <- long_all %>%
  group_by(.data[[QUESTION_COL]], time) %>%
  mutate(
    total = sum(count, na.rm = TRUE),
    pct   = count / total * 100
  ) %>%
  ungroup()

# ── Assign direction for diverging layout ─────────────────────────────────────
# Negative categories (disagree) extend left; positive extend right.
# Neutral is centred (split equally across the midline).
long_all <- long_all %>%
  mutate(
    direction = case_when(
      category %in% c("strongly_disagree", "disagree") ~ "negative",
      category %in% c("agree", "strongly_agree")        ~ "positive",
      TRUE                                               ~ "neutral"
    ),
    pct_directed = case_when(
      direction == "negative" ~ -pct,
      direction == "neutral"  ~ -pct / 2,   # half of neutral goes left
      TRUE                    ~  pct
    )
  )

# ── Colour scale (red → grey → blue, colourblind-friendly) ────────────────────
cat_colours <- c(
  strongly_disagree = "#d62728",
  disagree          = "#ff9896",
  neutral           = "#c7c7c7",
  agree             = "#aec7e8",
  strongly_agree    = "#1f77b4"
)[l_cat_stems]
cat_colours <- cat_colours[!is.na(cat_colours)]

ggplot(long_all,
       aes(x    = pct_directed,
           y    = interaction(.data[[QUESTION_COL]], time, sep = " – "),
           fill = category)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_vline(xintercept = 0, colour = "grey30", linewidth = 0.5) +
  scale_fill_manual(
    values = cat_colours,
    labels = str_to_title(str_replace_all(names(cat_colours), "_", " ")),
    guide  = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    labels = function(x) paste0(abs(round(x)), "%"),
    limits = c(-105, 105),
    breaks = seq(-100, 100, 25)
  ) +
  labs(
    title    = "Likert Questions: Response Distribution Pre vs. Post",
    subtitle = "Left = disagree · Right = agree · Neutral split at centre",
    x = NULL, y = NULL, fill = "Response"
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom",
        panel.grid.major.y = element_blank(),
        axis.text.y = element_text(size = 9))

} # end if (length(l_cat_stems) > 0)
```

## Visualisation — Average Score Change

```{r likert-plot-avgscores, fig.height=max(4, nrow(results_likert) * 0.55 + 1.5)}
# Lollipop chart: change in mean Likert score (Post − Pre) per question.
# Rendered only when average-score columns were detected in the data.
# Positive values indicate a more favourable response after the intervention.

if (is.null(pre_score_col)) {
  message("No avg_score columns detected — skipping average score plot.")
} else {
  ggplot(results_likert,
         aes(x = avg_diff,
             y = reorder(question, avg_diff),
             colour = significant)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
    geom_segment(aes(x = 0, xend = avg_diff, y = question, yend = question),
                 colour = "grey80", linewidth = 0.8) +
    geom_point(size = 4) +
    geom_text(
      aes(
        label = paste0(ifelse(avg_diff >= 0, "+", ""), avg_diff),
        hjust = ifelse(avg_diff >= 0, -0.4, 1.4)
      ),
      size = 3.2
    ) +
    scale_colour_manual(
      values   = c("TRUE" = "#e63946", "FALSE" = "#6c757d"),
      labels   = c("TRUE" = "Significant (p<.05)", "FALSE" = "Not significant"),
      na.value = "grey80",
      name     = NULL
    ) +
    scale_x_continuous(
      labels = function(x) paste0(ifelse(x >= 0, "+", ""), x),
      expand = expansion(mult = 0.25)
    ) +
    labs(
      title    = "Likert Questions: Change in Average Score (Post − Pre)",
      subtitle = paste0("Scale: ", LIKERT_MIN, "–", LIKERT_MAX,
                        " | Positive = more favourable post-training"),
      x = "Change in average score", y = NULL
    ) +
    theme_minimal(base_size = 12) +
    theme(legend.position = "top", panel.grid.major.y = element_blank())
}
```

---

# Multi-Select Questions Analysis

> **Statistical method:** Each response option is treated as a binary outcome
> (selected vs. not selected). A chi-square test compares the distribution of
> selection counts across all options at pre vs. post for each question.
> Per-option percentage tables show which specific options shifted most.

## Statistical Tests

```{r multiselect-stats}
# ══════════════════════════════════════════════════════════════════════════════
# Multi-Select: chi-square on option-count distribution
#
# For each question, we build a 2 × m contingency table (m = number of options):
#
#       | Opt A | Opt B | Opt C | … |
#   Pre |       |       |       |   |
#  Post |       |       |       |   |
#
# The chi-square test asks: did the overall pattern of which options respondents
# chose change between pre and post?
# Per-option % rows make it easy to identify which options drove any change.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Running Multi-Select analysis")

pre_m_cols  <- paste0("pre_",  m_option_stems)
post_m_cols <- paste0("post_", m_option_stems)

# Optional total-respondent columns (denominator for % calculation)
pre_m_total_col  <- if (!is.na(m_total_stem)) paste0("pre_",  m_total_stem) else NULL
post_m_total_col <- if (!is.na(m_total_stem)) paste0("post_", m_total_stem) else NULL

log_msg("Multi-Select — ", length(m_option_stems), " option stems | ",
        "total col: ", if(!is.null(pre_m_total_col)) pre_m_total_col else "none")

# ── Per-question chi-square + per-option table ────────────────────────────────
run_multiselect_test <- function(row) {
  q_label <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
               as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))

  pre_counts  <- safe_num(unlist(row[pre_m_cols]))
  post_counts <- safe_num(unlist(row[post_m_cols]))
  names(pre_counts)  <- m_option_stems
  names(post_counts) <- m_option_stems

  # Determine total respondents (used as denominator for percentages)
  pre_total  <- if (!is.null(pre_m_total_col))
    safe_num(row[[pre_m_total_col]]) else sum(pre_counts,  na.rm = TRUE)
  post_total <- if (!is.null(post_m_total_col))
    safe_num(row[[post_m_total_col]]) else sum(post_counts, na.rm = TRUE)

  valid <- !is.na(pre_counts) & !is.na(post_counts)
  mat   <- rbind(pre_counts[valid], post_counts[valid])

  if (sum(valid) < 2 || any(rowSums(mat) == 0))
    return(data.frame(
      question=q_label, option=NA_character_,
      pre_pct=NA_real_, post_pct=NA_real_, diff_pp=NA_real_,
      chi_sq=NA_real_, p_value=NA_real_, cramers_v=NA_real_,
      significant=NA, stringsAsFactors=FALSE
    ))

  test <- tryCatch(
    chisq.test(mat, simulate.p.value = any(mat < 5)),
    error = function(e) NULL
  )
  n  <- sum(mat)
  k  <- ncol(mat)
  cv <- if (!is.null(test)) round(cramers_v(test$statistic, n, k), 3) else NA_real_
  p  <- if (!is.null(test)) round(test$p.value, 4) else NA_real_
  cs <- if (!is.null(test)) round(test$statistic, 3) else NA_real_

  # Build one row per option (for granular visualisation)
  bind_rows(lapply(m_option_stems[valid], function(opt) {
    ptot <- if (is.na(pre_total)  || pre_total  == 0) NA_real_ else pre_total
    qtot <- if (is.na(post_total) || post_total == 0) NA_real_ else post_total
    data.frame(
      question    = q_label,
      option      = opt,
      pre_pct     = round(pre_counts[opt]  / ptot * 100, 1),
      post_pct    = round(post_counts[opt] / qtot * 100, 1),
      diff_pp     = round((post_counts[opt]/qtot - pre_counts[opt]/ptot) * 100, 1),
      chi_sq      = cs,
      p_value     = p,
      cramers_v   = cv,
      significant = if (!is.na(p)) p < ALPHA else NA,
      stringsAsFactors = FALSE
    )
  }))
}

if (length(m_option_stems) == 0) {
  log_msg("No option count columns detected in df_multiselect", level = "WARN")
  results_multiselect <- data.frame()
} else {
  results_multiselect <- bind_rows(
    lapply(seq_len(nrow(df_multiselect)),
           function(i) run_multiselect_test(df_multiselect[i, ]))
  )
  n_sig_m <- sum(results_multiselect$significant == TRUE, na.rm = TRUE)
  log_msg("Multi-Select: ", nrow(results_multiselect), " option rows | ",
          n_sig_m, " significant")

  results_multiselect %>%
    mutate(
      pre_pct_fmt  = paste0(pre_pct,  "%"),
      post_pct_fmt = paste0(post_pct, "%"),
      diff_fmt     = paste0(ifelse(diff_pp >= 0, "+", ""), diff_pp, "pp"),
      sig          = sig_flag(significant)
    ) %>%
    select(question, option, pre_pct_fmt, post_pct_fmt, diff_fmt,
           chi_sq, p_value, cramers_v, sig) %>%
    kable(
      col.names = c("Question", "Option", "Pre %", "Post %", "Change",
                    "χ²", "p-value", "Cramér's V", "Significant?"),
      caption   = "Multi-Select: chi-square test on option selection rates"
    ) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE) %>%
    row_spec(which(results_multiselect$significant == TRUE),
             bold = TRUE, color = "#e63946")
}
```

## Visualisation — Option Selection Rates

```{r multiselect-plot, fig.height=max(5, nrow(results_multiselect) * 0.45 + 2)}
# Grouped horizontal bar chart.
# Each bar represents the % of respondents who selected a given option.
# Options are labelled "Question — Option Name" for easy reading.
# Stars mark options belonging to questions with significant overall shifts.

if (nrow(results_multiselect) > 0) {
  plot_m <- results_multiselect %>%
    mutate(
      label_base = paste0(question, " — ",
                          str_to_title(str_replace_all(option, "_", " "))),
      label      = ifelse(!is.na(significant) & significant,
                          paste0(label_base, " *"), label_base)
    ) %>%
    pivot_longer(c(pre_pct, post_pct), names_to = "time", values_to = "pct") %>%
    mutate(
      time = recode(time, pre_pct = "Pre", post_pct = "Post"),
      time = factor(time, levels = c("Pre", "Post"))
    )

  ggplot(plot_m, aes(x = pct, y = reorder(label, pct), fill = time)) +
    geom_col(position = position_dodge(width = 0.7), width = 0.6, colour = "white") +
    geom_text(aes(label = paste0(round(pct, 1), "%")),
              position = position_dodge(width = 0.7), hjust = -0.15, size = 3) +
    scale_fill_manual(values = PAL, name = "Time point") +
    scale_x_continuous(labels = percent_format(scale = 1),
                       expand = expansion(mult = c(0, 0.2))) +
    labs(
      title    = "Multi-Select: % of Respondents Selecting Each Option",
      subtitle = "* indicates statistically significant distribution shift (p < 0.05)",
      x = "% Selected", y = NULL
    ) +
    theme_minimal(base_size = 11) +
    theme(legend.position = "top", panel.grid.major.y = element_blank(),
          axis.text.y = element_text(size = 9))
}
```

---

# Summary Table {#summary}

```{r summary-table}
# ══════════════════════════════════════════════════════════════════════════════
# SECTION 12 — COMBINED RESULTS SUMMARY
#
# All four question types are merged into a single sorted table.
# This is the at-a-glance view for stakeholders — it shows which questions
# changed significantly and by how much, regardless of question type.
# ══════════════════════════════════════════════════════════════════════════════

log_msg("Building combined summary table")

# ── Perception Slider ─────────────────────────────────────────────────────────
summary_s <- if (nrow(results_slider) > 0) {
  results_slider %>%
    transmute(
      question    = question,
      type        = "Perception Slider",
      metric      = paste0("mean: ", pre_mean, " → ", post_mean),
      change      = paste0(ifelse(diff >= 0 | grepl("^\\+", diff), "+", ""), diff),
      p_value     = p_value,
      effect_size = ifelse(!is.na(cohens_d), paste0("d = ", cohens_d), "—"),
      significant = significant
    )
} else data.frame()

# ── Knowledge ─────────────────────────────────────────────────────────────────
summary_k <- results_knowledge %>%
  transmute(
    question    = question,
    type        = "Knowledge",
    metric      = paste0("% correct: ", pre_pct, "% → ", post_pct, "%"),
    change      = paste0(ifelse(pct_point_diff >= 0, "+", ""),
                         pct_point_diff, "pp"),
    p_value     = p_value,
    effect_size = paste0(abs(pct_point_diff), "pp"),
    significant = significant
  )

# ── Likert ────────────────────────────────────────────────────────────────────
summary_l <- results_likert %>%
  transmute(
    question    = question,
    type        = "Likert",
    metric      = if (!is.null(pre_score_col))
                    paste0("avg score: ", pre_avg, " → ", post_avg)
                  else paste0("χ² = ", chi_sq),
    change      = if (!is.null(pre_score_col))
                    paste0(ifelse(!is.na(avg_diff) & avg_diff >= 0, "+", ""), avg_diff)
                  else NA_character_,
    p_value     = p_value,
    effect_size = ifelse(!is.na(cramers_v), paste0("V = ", cramers_v), "—"),
    significant = significant
  )

# ── Multi-Select (one row per question, summarised across options) ─────────────
summary_m <- if (nrow(results_multiselect) > 0) {
  results_multiselect %>%
    group_by(question) %>%
    summarise(
      n_sig  = sum(significant == TRUE, na.rm = TRUE),
      n_opts = n(),
      p_min  = min(p_value,  na.rm = TRUE),
      v_max  = max(cramers_v, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    transmute(
      question    = question,
      type        = "Multi-Select",
      metric      = paste0(n_sig, "/", n_opts, " options changed"),
      change      = NA_character_,
      p_value     = round(p_min, 4),
      effect_size = paste0("V = ", round(v_max, 3)),
      significant = p_min < ALPHA
    )
} else data.frame()

# ── Combine and render ─────────────────────────────────────────────────────────
combined_summary <- bind_rows(summary_s, summary_k, summary_l, summary_m) %>%
  arrange(type, p_value)

log_msg("Summary: ", nrow(combined_summary), " total items across all question types")

combined_summary %>%
  mutate(sig = sig_flag(significant)) %>%
  select(type, question, metric, change, p_value, effect_size, sig) %>%
  kable(
    col.names = c("Type", "Question", "Pre → Post", "Change",
                  "p-value", "Effect Size", "Sig."),
    caption   = paste(REPORT_TITLE, "| Complete Results Summary"),
    align     = "llllccl"
  ) %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = TRUE) %>%
  {
    tbl <- .
    type_order <- c("Perception Slider", "Knowledge", "Likert", "Multi-Select")
    for (tp in type_order) {
      rows <- which(combined_summary$type == tp)
      if (length(rows) > 0)
        tbl <- pack_rows(tbl, tp, min(rows), max(rows), bold = TRUE)
    }
    tbl
  } %>%
  row_spec(which(combined_summary$significant == TRUE),
           bold = TRUE, color = "#e63946")
```
---
---

# Knowledge Questions (Improved)

```{r knowledge-stats-v2}
log_msg("Running improved Knowledge analysis")

kcols <- list(
  pre_correct  = "pre_correct",
  post_correct = "post_correct",
  pre_total    = "pre_total",
  post_total   = "post_total"
)

run_prop_test_v2 <- function(row, cols) {
  q_label <- as.character(row[["question_text"]])
  x1 <- as.numeric(row[[cols$pre_correct]]); n1 <- as.numeric(row[[cols$pre_total]])
  x2 <- as.numeric(row[[cols$post_correct]]); n2 <- as.numeric(row[[cols$post_total]])

  test <- prop.test(c(x1, x2), c(n1, n2), correct = FALSE)

  data.frame(
    question_text  = q_label,
    pre_pct        = round(x1/n1*100, 1),
    post_pct       = round(x2/n2*100, 1),
    pct_point_diff = round((x2/n2 - x1/n1)*100, 1),
    # flip sign: prop.test gives pre-post, we want post-pre
    ci_low         = round(-test$conf.int[2]*100, 1),
    ci_high        = round(-test$conf.int[1]*100, 1),
    p_value        = round(test$p.value, 4),
    significant    = test$p.value < 0.05,
    stringsAsFactors = FALSE
  )
}

results_knowledge_v2 <- bind_rows(
  lapply(seq_len(nrow(df_knowledge)),
         function(i) run_prop_test_v2(df_knowledge[i, ], kcols))
)

log_msg("Knowledge v2: ", nrow(results_knowledge_v2), " questions | ",
        sum(results_knowledge_v2$significant), " significant")

results_knowledge_v2 %>%
  mutate(
    pre_pct_fmt  = paste0(pre_pct, "%"),
    post_pct_fmt = paste0(post_pct, "%"),
    change_fmt   = paste0(ifelse(pct_point_diff >= 0, "+", ""), pct_point_diff,
                          "pp  [", ci_low, ", +", ci_high, "]"),
    sig          = sig_flag(significant)
  ) %>%
  select(question_text, pre_pct_fmt, post_pct_fmt, change_fmt, p_value, sig) %>%
  kable(
    col.names = c("Question", "Pre %", "Post %", "Change (95% CI)", "p-value", "Significant?"),
    caption   = "Two-proportion z-test: pre vs. post % correct (improved)"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(which(results_knowledge_v2$significant == TRUE),
           bold = TRUE, color = "#e63946")
```
```

## Visualisation — % Correct Pre vs. Post

```{r knowledge-plot-bar-v2, fig.height=max(4, nrow(results_knowledge_v2) * 0.65 + 1.5)}
PAL_v2 <- c(Pre = "#0077b6", Post = "#f4a261")

plot_k2 <- results_knowledge_v2 %>%
  filter(!is.na(pre_pct), !is.na(post_pct)) %>%
  pivot_longer(c(pre_pct, post_pct), names_to = "time", values_to = "pct") %>%
  mutate(
    time = recode(time, pre_pct = "Pre", post_pct = "Post"),
    time = factor(time, levels = c("Pre", "Post"))
  )

sig_badges <- results_knowledge_v2 %>%
  filter(!is.na(significant), significant == TRUE) %>%
  mutate(
    x_nudge = pmax(pre_pct, post_pct) + 5,
    badge   = paste0("+", pct_point_diff, "pp \u2713")
  )

ggplot(plot_k2,
       aes(x = pct, y = reorder(question_text, pct), fill = time)) +
  geom_col(position = position_dodge(width = 0.72), width = 0.65,
           colour = "white", linewidth = 0.3) +
  geom_text(aes(label = paste0(round(pct, 1), "%")),
            position = position_dodge(width = 0.72),
            hjust = -0.2, size = 3.1, colour = "grey40") +
  geom_text(data = sig_badges,
            aes(x = x_nudge, y = question_text, label = badge),
            inherit.aes = FALSE,
            hjust = 0, size = 3.2, colour = "#993C1D", fontface = "bold") +
  scale_fill_manual(values = PAL_v2, name = NULL) +
  scale_x_continuous(labels = percent_format(scale = 1),
                     expand = expansion(mult = c(0, 0.3))) +
  labs(
    title    = "Knowledge questions: % correct responses",
    subtitle = "Sorted by post score \u00b7 Red labels = significant change (p < 0.05)",
    x = "% correct", y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position    = "top",
        panel.grid.major.y = element_blank(),
        axis.text.y        = element_text(size = 10))
```

## Visualisation — Change with Confidence Intervals

```{r knowledge-plot-dot-v2, fig.height=max(4, nrow(results_knowledge_v2) * 0.65 + 1.5)}
plot_ci2 <- results_knowledge_v2 %>%
  filter(!is.na(pct_point_diff)) %>%
  mutate(
    sig_label = case_when(
      is.na(significant)  ~ "No data",
      significant == TRUE ~ "Significant (p \u2264 .05)",
      TRUE                ~ "Not significant"
    )
  )

ggplot(plot_ci2,
       aes(x = pct_point_diff,
           y = reorder(question_text, pct_point_diff),
           colour = sig_label)) +
  geom_vline(xintercept = 0, linetype = "dashed",
             colour = "grey50", linewidth = 0.6) +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high),
                 height = 0.25, linewidth = 0.8) +
  geom_point(size = 4.5) +
  # label sits ABOVE the point, no horizontal overlap with CI bars
  geom_text(
    aes(label = paste0(ifelse(pct_point_diff >= 0, "+", ""), pct_point_diff, "pp")),
    vjust = -1.1,
    hjust = 0.5,
    size = 3.2
  ) +
  scale_colour_manual(
    values = c(
      "Significant (p \u2264 .05)" = "#993C1D",
      "Not significant"            = "#6c757d",
      "No data"                    = "grey85"
    ),
    name = NULL
  ) +
  # set x limits explicitly so the axis doesn't run past the data
  scale_x_continuous(
    labels = function(x) paste0(ifelse(x >= 0, "+", ""), x, "pp"),
    limits = c(-10, 65),
    expand = expansion(mult = 0.05)
  ) +
  labs(
    title    = "Knowledge questions: change with 95% confidence intervals",
    subtitle = "Error bars = 95% CI \u00b7 Sorted by effect size\nCI crossing zero = not significant",
    x = "Percentage-point change (post \u2212 pre)", y = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position    = "top",
        panel.grid.major.y = element_blank(),
        axis.text.y        = element_text(size = 10))
```
---

# Session Info

```{r session-info, echo=FALSE}
# ══════════════════════════════════════════════════════════════════════════════
# REPRODUCIBILITY RECORD
#
# The output below records the exact R version and package versions used
# when this report was generated. If results ever differ between runs,
# this information helps diagnose version-related causes.
# ══════════════════════════════════════════════════════════════════════════════
log_msg("Report complete — recording session info")
sessionInfo()
```
