---
title: "Pre–Post Survey Analysis"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    df_print: paged
  word_document:
    toc: true
    toc_depth: '3'
  pdf_document:
    toc: true
    toc_depth: '3'
---

```{css, echo=FALSE}
/* Minor style improvements */
h1.title { font-size: 2rem; font-weight: 700; }
.section-label { color: #2c7be5; font-weight: 600; }
table { font-size: 0.9rem; }
.sig   { color: #e63946; font-weight: bold; }
.nosig { color: #6c757d; }
```

---

> **How to use this document**
> 1. Fill in the **Configuration** block in Section 1.
> 2. Run `superset_connection.R` once to pull data from Superset (or place cached CSVs in `data_cache/`).
> 3. Click **Knit** — the report re-generates automatically every time.
> Column names are detected automatically from your data; no other edits are needed.

---

# Setup & Configuration {.tabset}

## Libraries

```{r setup, message=FALSE, warning=FALSE}
# ── Install missing packages on first run ──────────────────────────────────────
required_pkgs <- c(
  "dplyr", "tidyr", "ggplot2", "scales",
  "knitr", "kableExtra", "purrr", "stringr",
  "rstatix", "effectsize"
)
new_pkgs <- required_pkgs[!sapply(required_pkgs, requireNamespace, quietly = TRUE)]
if (length(new_pkgs) > 0) install.packages(new_pkgs, repos = "https://cloud.r-project.org")
invisible(lapply(required_pkgs, library, character.only = TRUE))

knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 9,
  fig.height = 5.5,
  dpi        = 150
)

# Consistent colour palette (pre = teal, post = orange)
PAL <- c(pre = "#0077b6", post = "#f4a261")

# Significance threshold
ALPHA <- 0.05
```

## Configuration

```{r config}
# ── EDIT THESE VALUES ──────────────────────────────────────────────────────────

# Name of the column that identifies each survey question (row label)
# e.g., "question", "item", "question_text"
QUESTION_COL <- "question"   # <-- update to match your actual column name

# Optional: a character vector of question IDs/labels to EXCLUDE from analysis
# e.g., c("Q99", "demographic_item")
EXCLUDE_QUESTIONS <- c()

# Optional: custom question label column for prettier axis labels
# Set to NULL to use QUESTION_COL as-is
LABEL_COL <- NULL   # e.g., "question_text"

# Likert scale range used in your survey (affects avg-score comparisons)
LIKERT_MIN <- 1
LIKERT_MAX <- 5   # change to 7 if using a 7-point scale

# Report title (appears in summary table header)
REPORT_TITLE <- "Pre–Post Survey Results"

# ── END EDITABLE SECTION ───────────────────────────────────────────────────────
```

---

# Data Import

```{r import-data}
# ── Load data ──────────────────────────────────────────────────────────────────
# Option A (recommended): source superset_connection.R to pull live data.
# Option B: load from CSVs cached by superset_connection.R (offline/reproducible).
#
# Uncomment ONE of the two options:

# source("superset_connection.R")              # Option A – live Superset pull

# Option B – load from local cache (run superset_connection.R at least once first)
df_slider      <- read.csv("data_cache/slider.csv",
                            stringsAsFactors = FALSE, check.names = FALSE)
df_knowledge   <- read.csv("data_cache/knowledge.csv",
                            stringsAsFactors = FALSE, check.names = FALSE)
df_likert      <- read.csv("data_cache/likert.csv",
                            stringsAsFactors = FALSE, check.names = FALSE)
df_multiselect <- read.csv("data_cache/multiselect.csv",
                            stringsAsFactors = FALSE, check.names = FALSE)

# ── Auto-detect QUESTION_COL if the configured value isn't present ─────────────
detect_q_col <- function(df) {
  # Check common question-label column names first
  candidates <- c("question", "Question", "question_text", "Question Text",
                  "item", "Item", "label", "Label", "survey_question",
                  "row_label", "name", "Name", "description")
  found <- intersect(candidates, names(df))
  if (length(found) > 0) return(found[1])
  # Fall back to first non-pre_/post_ column (likely the row identifier)
  other <- names(df)[!grepl("^(pre_|post_)", names(df))]
  if (length(other) > 0) return(other[1])
  return(names(df)[1])   # last resort: very first column
}

if (!QUESTION_COL %in% names(df_knowledge)) {
  detected <- detect_q_col(df_knowledge)
  message("⚠  QUESTION_COL '", QUESTION_COL, "' not found in data. ",
          "Auto-detected: '", detected, "'  — update QUESTION_COL in the ",
          "Configuration block to silence this message.")
  QUESTION_COL <- detected
}
cat("✔ Using QUESTION_COL:", QUESTION_COL, "\n")

# ── Sanity checks ──────────────────────────────────────────────────────────────
for (nm in c("df_knowledge", "df_likert", "df_slider", "df_multiselect")) {
  df <- get(nm)
  if (!QUESTION_COL %in% names(df)) {
    warning(nm, ": QUESTION_COL '", QUESTION_COL, "' not found. ",
            "Available columns: ", paste(names(df), collapse = ", "))
  }
}

# ── Apply exclusions ───────────────────────────────────────────────────────────
exclude_rows <- function(df) {
  if (length(EXCLUDE_QUESTIONS) > 0 && QUESTION_COL %in% names(df))
    df[!df[[QUESTION_COL]] %in% EXCLUDE_QUESTIONS, ]
  else df
}
df_slider      <- exclude_rows(df_slider)
df_knowledge   <- exclude_rows(df_knowledge)
df_likert      <- exclude_rows(df_likert)
df_multiselect <- exclude_rows(df_multiselect)

cat(sprintf("%-16s : %d rows x %d cols\n", "df_slider",      nrow(df_slider),      ncol(df_slider)))
cat(sprintf("%-16s : %d rows x %d cols\n", "df_knowledge",   nrow(df_knowledge),   ncol(df_knowledge)))
cat(sprintf("%-16s : %d rows x %d cols\n", "df_likert",      nrow(df_likert),      ncol(df_likert)))
cat(sprintf("%-16s : %d rows x %d cols\n", "df_multiselect", nrow(df_multiselect), ncol(df_multiselect)))
```

## Raw data preview

```{r data-preview}
cat("── Perception Slider ──\n");  head(df_slider)
cat("── Knowledge ──\n");          head(df_knowledge)
cat("── Likert ──\n");             head(df_likert)
cat("── Multi-Select ──\n");       head(df_multiselect)
```

---

# Column Auto-Detection

```{r detect-columns}
# ── Automatically identify pre_ and post_ column pairs for each dataset ────────
detect_pairs <- function(df) {
  pre_cols  <- grep("^pre_",  names(df), value = TRUE)
  post_cols <- grep("^post_", names(df), value = TRUE)
  pre_stems  <- sub("^pre_",  "", pre_cols)
  post_stems <- sub("^post_", "", post_cols)
  matched    <- intersect(pre_stems, post_stems)
  list(
    pre      = paste0("pre_",  matched),
    post     = paste0("post_", matched),
    stems    = matched,
    pre_only = setdiff(pre_stems,  post_stems),
    post_only= setdiff(post_stems, pre_stems)
  )
}

pairs_s <- detect_pairs(df_slider)
pairs_k <- detect_pairs(df_knowledge)
pairs_l <- detect_pairs(df_likert)
pairs_m <- detect_pairs(df_multiselect)

for (nm in c("Slider","Knowledge","Likert","Multi-Select")) {
  p <- list(Slider=pairs_s, Knowledge=pairs_k,
             Likert=pairs_l, `Multi-Select`=pairs_m)[[nm]]
  cat(nm, "– matched pairs:", paste(p$stems, collapse=", "), "\n")
  if (length(p$pre_only)  > 0)
    cat("  pre-only  :", paste(p$pre_only,  collapse=", "), "\n")
  if (length(p$post_only) > 0)
    cat("  post-only :", paste(p$post_only, collapse=", "), "\n")
}
```

```{r detect-column-types}
# ── Classify column types within each dataset ─────────────────────────────────

# Slider: detect numeric/score stems (avg, mean, score, value)
s_score_stems <- pairs_s$stems[grepl("avg|mean|score|value",
                                      pairs_s$stems, ignore.case = TRUE)]
s_count_stems <- pairs_s$stems[!grepl("avg|mean|score|value|pct|percent|delta",
                                       pairs_s$stems, ignore.case = TRUE)]
cat("Slider score stems     :", paste(s_score_stems, collapse = ", "), "\n")
cat("Slider count stems     :", paste(s_count_stems, collapse = ", "), "\n")

# Knowledge: count, pct, and delta cols
k_count_stems <- pairs_k$stems[!grepl("pct|percent|delta|avg|mean|score",
                                       pairs_k$stems, ignore.case = TRUE)]
k_pct_stems   <- pairs_k$stems[ grepl("pct|percent",
                                       pairs_k$stems, ignore.case = TRUE)]
k_delta_cols  <- grep("delta", names(df_knowledge), value = TRUE)
cat("\nKnowledge count stems  :", paste(k_count_stems, collapse = ", "), "\n")
cat("Knowledge pct stems    :", paste(k_pct_stems,   collapse = ", "), "\n")
cat("Knowledge delta cols   :", paste(k_delta_cols,  collapse = ", "), "\n")

# Likert: category cols vs avg_score
likert_cat_order <- c("strongly_disagree", "disagree", "neutral",
                      "agree", "strongly_agree")
l_cat_stems  <- intersect(likert_cat_order, pairs_l$stems)
l_score_stem <- pairs_l$stems[grepl("avg|mean|score", pairs_l$stems,
                                     ignore.case = TRUE)]
cat("\nLikert category stems  :", paste(l_cat_stems,  collapse = ", "), "\n")
cat("Likert score stems     :", paste(l_score_stem,  collapse = ", "), "\n")

# Multi-Select: all paired stems are option count columns
m_option_stems <- pairs_m$stems[!grepl("pct|percent|delta|avg|mean|total",
                                        pairs_m$stems, ignore.case = TRUE)]
m_total_stem   <- pairs_m$stems[grepl("total", pairs_m$stems, ignore.case=TRUE)][1]
cat("\nMulti-Select option stems:", paste(m_option_stems, collapse = ", "), "\n")
cat("Multi-Select total stem  :", m_total_stem, "\n")
```

---

# Perception Slider Analysis

> **Method:** Descriptive comparison of pre vs. post mean scores per question.
> Because only aggregate means are available (not individual responses), formal
> inferential testing requires a standard deviation and sample size. If those
> columns exist in your data they are used for a two-sample t-test; otherwise
> the analysis is descriptive with effect size reported as Cohen's d estimate.

## Statistical Tests

```{r slider-stats}
# Auto-detect pre/post score columns
pre_s_col  <- if (length(s_score_stems) > 0) paste0("pre_",  s_score_stems[1]) else NULL
post_s_col <- if (length(s_score_stems) > 0) paste0("post_", s_score_stems[1]) else NULL

# Try to detect n, sd columns (used for proper t-test if available)
pre_n_col   <- grep("^pre_(n|count|total|respondents)$",
                    names(df_slider), value=TRUE, ignore.case=TRUE)[1]
post_n_col  <- grep("^post_(n|count|total|respondents)$",
                    names(df_slider), value=TRUE, ignore.case=TRUE)[1]
pre_sd_col  <- grep("^pre_(sd|std|stdev)",
                    names(df_slider), value=TRUE, ignore.case=TRUE)[1]
post_sd_col <- grep("^post_(sd|std|stdev)",
                    names(df_slider), value=TRUE, ignore.case=TRUE)[1]

has_ttest_inputs <- !is.na(pre_n_col) && !is.na(post_n_col) &&
                    !is.na(pre_sd_col) && !is.na(post_sd_col)

run_slider_test <- function(row) {
  q_label   <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
                 as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))
  pre_mean  <- suppressWarnings(as.numeric(row[[pre_s_col]]))
  post_mean <- suppressWarnings(as.numeric(row[[post_s_col]]))
  diff      <- round(post_mean - pre_mean, 3)

  if (has_ttest_inputs) {
    n1  <- as.numeric(row[[pre_n_col]]);  n2  <- as.numeric(row[[post_n_col]])
    sd1 <- as.numeric(row[[pre_sd_col]]); sd2 <- as.numeric(row[[post_sd_col]])
    # Welch two-sample t-test from summary statistics
    se   <- sqrt(sd1^2/n1 + sd2^2/n2)
    t    <- diff / se
    df_t <- (sd1^2/n1 + sd2^2/n2)^2 /
            ((sd1^2/n1)^2/(n1-1) + (sd2^2/n2)^2/(n2-1))
    p    <- 2 * pt(-abs(t), df_t)
    # Cohen's d (pooled)
    sp   <- sqrt(((n1-1)*sd1^2 + (n2-1)*sd2^2) / (n1+n2-2))
    d    <- round(diff / sp, 3)
    data.frame(question=q_label,
               pre_mean=round(pre_mean,2), post_mean=round(post_mean,2),
               diff=diff, p_value=round(p,4), cohens_d=d,
               significant=p < ALPHA, method="Welch t-test")
  } else {
    # Descriptive only: Cohen's d not calculable without SD
    data.frame(question=q_label,
               pre_mean=round(pre_mean,2), post_mean=round(post_mean,2),
               diff=diff, p_value=NA, cohens_d=NA,
               significant=NA, method="Descriptive (no SD available)")
  }
}

if (is.null(pre_s_col)) {
  message("No score columns detected in df_slider. Check column names.")
  results_slider <- data.frame()
} else {
  results_slider <- bind_rows(
    lapply(seq_len(nrow(df_slider)), function(i) run_slider_test(df_slider[i,]))
  )

  results_slider %>%
    mutate(
      diff      = paste0(ifelse(diff >= 0, "+", ""), diff),
      sig_flag  = case_when(is.na(significant) ~ "—",
                            significant        ~ "✓ Yes",
                            TRUE               ~ "No")
    ) %>%
    select(question, pre_mean, post_mean, diff, p_value, cohens_d, sig_flag, method) %>%
    kable(
      col.names = c("Question","Pre Mean","Post Mean","Change",
                    "p-value","Cohen's d","Significant?","Method"),
      caption   = "Perception Slider: Pre vs. Post Mean Score Comparison"
    ) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE) %>%
    row_spec(which(results_slider$significant == TRUE),
             bold = TRUE, color = "#e63946")
}
```

## Visualisation — Slider Score Change

```{r slider-plot, fig.height=max(4, nrow(df_slider) * 0.55 + 1.5)}
if (nrow(results_slider) > 0) {
  # Long format for grouped bar
  plot_s <- results_slider %>%
    select(question, pre_mean, post_mean) %>%
    pivot_longer(c(pre_mean, post_mean), names_to="time", values_to="score") %>%
    mutate(time = recode(time, pre_mean="Pre", post_mean="Post"),
           time = factor(time, levels=c("Pre","Post")))

  sig_lbl_s <- results_slider %>%
    mutate(label = ifelse(!is.na(significant) & significant,
                          paste0(question," *"), question)) %>%
    select(question, label)
  plot_s <- left_join(plot_s, sig_lbl_s, by="question")

  ggplot(plot_s, aes(x=score, y=reorder(label, score), fill=time)) +
    geom_col(position=position_dodge(width=0.7), width=0.6, colour="white") +
    geom_text(aes(label=round(score,2)),
              position=position_dodge(width=0.7), hjust=-0.15, size=3.2) +
    scale_fill_manual(values=PAL, name="Time point") +
    scale_x_continuous(expand=expansion(mult=c(0,0.2))) +
    labs(title="Perception Slider: Mean Score Pre vs. Post",
         subtitle="* statistically significant change (p < 0.05)",
         x="Mean Score", y=NULL) +
    theme_minimal(base_size=12) +
    theme(legend.position="top", panel.grid.major.y=element_blank())
}
```

---

# Knowledge Questions Analysis

> **Method:** Two-proportion z-test (`prop.test`) comparing the proportion of
> correct responses at pre vs. post for each question. Requires `pre_correct`,
> `pre_total`, `post_correct`, and `post_total` columns (or equivalent detected
> columns). Effect size reported as absolute percentage-point change.

## Statistical Tests

```{r knowledge-stats}
# ── Identify the correct/total count columns ───────────────────────────────────
# Looks for columns whose stem contains "correct" and "total"
correct_stem <- k_count_stems[grepl("correct",  k_count_stems, ignore.case = TRUE)][1]
total_stem   <- k_count_stems[grepl("total",    k_count_stems, ignore.case = TRUE)][1]

if (is.na(correct_stem) || is.na(total_stem)) {
  stop("Could not auto-detect 'correct' and 'total' count columns in df_knowledge. ",
       "Check that these columns follow the pre_/post_ naming pattern.")
}

pre_correct_col  <- paste0("pre_",  correct_stem)
post_correct_col <- paste0("post_", correct_stem)
pre_total_col    <- paste0("pre_",  total_stem)
post_total_col   <- paste0("post_", total_stem)

cat("Using columns:\n",
    " pre correct :", pre_correct_col,  "\n",
    " post correct:", post_correct_col, "\n",
    " pre total   :", pre_total_col,    "\n",
    " post total  :", post_total_col,   "\n")

# ── Run prop.test for each question ────────────────────────────────────────────
run_prop_test <- function(row) {
  q_label <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
               as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))

  x1 <- as.numeric(row[[pre_correct_col]])
  n1 <- as.numeric(row[[pre_total_col]])
  x2 <- as.numeric(row[[post_correct_col]])
  n2 <- as.numeric(row[[post_total_col]])

  # Skip if any value is NA or total is 0
  if (any(is.na(c(x1, n1, x2, n2))) || n1 == 0 || n2 == 0)
    return(data.frame(
      question       = q_label,
      pre_pct        = NA, post_pct = NA,
      pct_point_diff = NA, p_value  = NA,
      significant    = NA, note     = "Missing data"
    ))

  test <- tryCatch(
    prop.test(c(x1, x2), c(n1, n2), correct = FALSE),
    error = function(e) NULL
  )

  data.frame(
    question       = q_label,
    pre_pct        = round(x1 / n1 * 100, 1),
    post_pct       = round(x2 / n2 * 100, 1),
    pct_point_diff = round((x2 / n2 - x1 / n1) * 100, 1),
    p_value        = if (!is.null(test)) round(test$p.value, 4) else NA,
    significant    = if (!is.null(test)) test$p.value < ALPHA else NA,
    note           = ""
  )
}

results_knowledge <- bind_rows(
  lapply(seq_len(nrow(df_knowledge)), function(i) run_prop_test(df_knowledge[i, ]))
)

# Display results table
results_knowledge %>%
  mutate(
    pre_pct        = paste0(pre_pct, "%"),
    post_pct       = paste0(post_pct, "%"),
    pct_point_diff = paste0(ifelse(pct_point_diff >= 0, "+", ""), pct_point_diff, "pp"),
    sig_flag       = ifelse(significant, "✓ Yes", "No")
  ) %>%
  select(question, pre_pct, post_pct, pct_point_diff, p_value, sig_flag) %>%
  kable(
    col.names = c("Question", "Pre %", "Post %", "Change", "p-value", "Significant?"),
    caption   = "Two-proportion z-test: pre vs. post % correct",
    align     = "lccccc"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  row_spec(
    which(results_knowledge$significant == TRUE),
    bold = TRUE, color = "#e63946"
  )
```

## Visualisation — % Correct Pre vs Post

```{r knowledge-plot-bar, fig.height=max(4, nrow(results_knowledge) * 0.55 + 1.5)}
# Reshape for grouped bar chart
plot_k <- results_knowledge %>%
  mutate(pre_pct  = as.numeric(sub("%", "", pre_pct)),
         post_pct = as.numeric(sub("%", "", post_pct))) %>%
  select(question, pre_pct, post_pct) %>%
  pivot_longer(cols = c(pre_pct, post_pct),
               names_to  = "time",
               values_to = "pct_correct") %>%
  mutate(time = recode(time, pre_pct = "Pre", post_pct = "Post"),
         time = factor(time, levels = c("Pre", "Post")))

# Add significance markers to question labels
sig_labels <- results_knowledge %>%
  mutate(label = ifelse(significant == TRUE,
                        paste0(question, " *"),
                        question)) %>%
  select(question, label)

plot_k <- left_join(plot_k, sig_labels, by = "question")

ggplot(plot_k, aes(x = pct_correct, y = reorder(label, pct_correct),
                   fill = time)) +
  geom_col(position = position_dodge(width = 0.7),
           width = 0.6, colour = "white") +
  geom_text(aes(label = paste0(round(pct_correct, 1), "%")),
            position = position_dodge(width = 0.7),
            hjust = -0.15, size = 3.2) +
  scale_fill_manual(values = PAL, name = "Time point") +
  scale_x_continuous(labels = percent_format(scale = 1),
                     expand = expansion(mult = c(0, 0.18))) +
  labs(
    title    = "Knowledge Questions: % Correct Responses",
    subtitle = "* indicates statistically significant change (p < 0.05)",
    x        = "% Correct",
    y        = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position  = "top",
    panel.grid.major.y = element_blank(),
    axis.text.y      = element_text(size = 10)
  )
```

## Visualisation — Change in % Correct (Dot-Plot)

```{r knowledge-plot-dot, fig.height=max(4, nrow(results_knowledge) * 0.55 + 1.5)}
ggplot(results_knowledge,
       aes(x = pct_point_diff,
           y = reorder(question, pct_point_diff),
           colour = significant)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_segment(aes(x = 0, xend = pct_point_diff,
                   y = question, yend = question),
               colour = "grey80", linewidth = 0.8) +
  geom_point(size = 4) +
  geom_text(aes(label = paste0(ifelse(pct_point_diff >= 0, "+", ""),
                               pct_point_diff, "pp")),
            hjust = ifelse(results_knowledge$pct_point_diff >= 0, -0.3, 1.3),
            size  = 3.2) +
  scale_colour_manual(
    values = c("TRUE" = "#e63946", "FALSE" = "#6c757d"),
    labels = c("TRUE" = "Significant (p<.05)", "FALSE" = "Not significant"),
    name   = NULL
  ) +
  scale_x_continuous(
    labels = function(x) paste0(ifelse(x >= 0, "+", ""), x, "pp"),
    expand = expansion(mult = 0.2)
  ) +
  labs(
    title    = "Knowledge Questions: Percentage-Point Change (Post − Pre)",
    subtitle = "Positive = improvement; red = statistically significant",
    x        = "Percentage-point change",
    y        = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position    = "top",
    panel.grid.major.y = element_blank()
  )
```

---

# Likert Questions Analysis

> **Method:** Chi-square test comparing the frequency distribution of responses
> across Likert categories at pre vs. post for each question. Average score
> comparisons are descriptive only (no individual-level data available for a
> proper paired test). Effect size reported as Cramér's V.

## Statistical Tests

```{r likert-stats}
# ── Build column name vectors from detected stems ─────────────────────────────
pre_cat_cols  <- paste0("pre_",  l_cat_stems)
post_cat_cols <- paste0("post_", l_cat_stems)
pre_score_col  <- if (length(l_score_stem) > 0) paste0("pre_",  l_score_stem[1]) else NULL
post_score_col <- if (length(l_score_stem) > 0) paste0("post_", l_score_stem[1]) else NULL

cat("Category columns used in chi-square:\n",
    "  Pre :", paste(pre_cat_cols,  collapse = ", "), "\n",
    "  Post:", paste(post_cat_cols, collapse = ", "), "\n")
if (!is.null(pre_score_col))
  cat("Score columns:", pre_score_col, "→", post_score_col, "\n")

# ── Helper: Cramér's V ─────────────────────────────────────────────────────────
cramers_v <- function(chi2, n, k) {
  sqrt(chi2 / (n * (k - 1)))
}

# ── Run chi-square for each Likert question ────────────────────────────────────
run_chisq <- function(row) {
  q_label <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
               as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))
  pre_counts  <- suppressWarnings(
    as.numeric(row[pre_cat_cols]))
  post_counts <- suppressWarnings(
    as.numeric(row[post_cat_cols]))

  # Remove NA categories (columns present but empty for this question)
  valid <- !is.na(pre_counts) & !is.na(post_counts)
  pre_counts  <- pre_counts[valid]
  post_counts <- post_counts[valid]

  if (length(pre_counts) < 2 ||
      sum(pre_counts) == 0 || sum(post_counts) == 0) {
    return(data.frame(
      question    = q_label,
      pre_avg     = NA, post_avg = NA, avg_diff = NA,
      chi_sq      = NA, df = NA, p_value  = NA,
      cramers_v   = NA, significant = NA,
      note        = "Insufficient data"
    ))
  }

  mat  <- rbind(pre_counts, post_counts)
  test <- tryCatch(chisq.test(mat, simulate.p.value = any(mat < 5)),
                   error = function(e) NULL)

  n_pre  <- sum(pre_counts)
  n_post <- sum(post_counts)
  k      <- length(pre_counts)

  pre_avg  <- if (!is.null(pre_score_col))
    suppressWarnings(as.numeric(row[[pre_score_col]]))  else NA
  post_avg <- if (!is.null(post_score_col))
    suppressWarnings(as.numeric(row[[post_score_col]])) else NA

  data.frame(
    question    = q_label,
    pre_avg     = round(pre_avg,  2),
    post_avg    = round(post_avg, 2),
    avg_diff    = round(post_avg - pre_avg, 2),
    chi_sq      = if (!is.null(test)) round(test$statistic, 3) else NA,
    df          = if (!is.null(test)) test$parameter else NA,
    p_value     = if (!is.null(test)) round(test$p.value, 4)  else NA,
    cramers_v   = if (!is.null(test))
      round(cramers_v(test$statistic,
                      n_pre + n_post, k), 3) else NA,
    significant = if (!is.null(test)) test$p.value < ALPHA else NA,
    note        = if (any(mat < 5)) "Simulated p-value (low counts)" else ""
  )
}

results_likert <- bind_rows(
  lapply(seq_len(nrow(df_likert)), function(i) run_chisq(df_likert[i, ]))
)

# Display results table
results_likert %>%
  mutate(
    avg_diff = paste0(ifelse(!is.na(avg_diff) & avg_diff >= 0, "+", ""), avg_diff),
    sig_flag = ifelse(significant, "✓ Yes", "No")
  ) %>%
  select(question, pre_avg, post_avg, avg_diff,
         chi_sq, df, p_value, cramers_v, sig_flag) %>%
  kable(
    col.names = c("Question", "Pre Avg", "Post Avg", "Δ Avg",
                  "χ²", "df", "p-value", "Cramér's V", "Significant?"),
    caption   = paste0("Chi-square test: distribution of Likert responses pre vs. post\n",
                       "(Cramér's V: 0.10 = small, 0.30 = medium, 0.50 = large effect)"),
    align     = "lcccccccc"
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = FALSE
  ) %>%
  row_spec(
    which(results_likert$significant == TRUE),
    bold = TRUE, color = "#e63946"
  ) %>%
  footnote(
    general = results_likert$note[results_likert$note != ""],
    general_title = "Notes: "
  )
```

## Visualisation — Diverging Stacked Bar Chart

```{r likert-plot-diverging, fig.height=max(5, nrow(df_likert) * 1.1 + 2)}
# ── Build long-format data for diverging bars ─────────────────────────────────
build_diverging <- function(df, time_label, cat_cols, cat_stems) {

  df %>%
    select(all_of(c(QUESTION_COL, cat_cols))) %>%
    pivot_longer(
      cols      = all_of(cat_cols),
      names_to  = "category",
      values_to = "count"
    ) %>%
    mutate(
      category = str_replace(category, paste0("^", tolower(time_label), "_"), ""),
      time     = time_label,
      count    = as.numeric(count)
    )
}

long_pre  <- build_diverging(df_likert, "pre",  pre_cat_cols,  l_cat_stems)
long_post <- build_diverging(df_likert, "post", post_cat_cols, l_cat_stems)
long_all  <- bind_rows(long_pre, long_post) %>%
  mutate(
    time     = factor(time, levels = c("pre", "post"),
                      labels = c("Pre", "Post")),
    category = factor(category,
                      levels = rev(likert_cat_order[likert_cat_order %in% l_cat_stems]))
  )

# Convert to row percentages within each question × time group
long_all <- long_all %>%
  group_by(.data[[QUESTION_COL]], time) %>%
  mutate(
    total = sum(count, na.rm = TRUE),
    pct   = count / total * 100
  ) %>%
  ungroup()

# Define direction: negative categories shift left, positive shift right
negative_cats <- c("strongly_disagree", "disagree")
positive_cats <- c("agree", "strongly_agree")
neutral_cat   <- "neutral"

long_all <- long_all %>%
  mutate(
    direction = case_when(
      category %in% negative_cats ~ "negative",
      category %in% positive_cats ~ "positive",
      TRUE                         ~ "neutral"
    ),
    pct_directed = case_when(
      direction == "negative" ~ -pct,
      direction == "neutral"  ~  pct / 2 * -1,   # split neutral down the middle
      TRUE                    ~  pct
    )
  )

# Colour scale (red → orange → grey → light blue → dark blue)
cat_colours <- c(
  strongly_disagree = "#d62728",
  disagree          = "#ff9896",
  neutral           = "#c7c7c7",
  agree             = "#aec7e8",
  strongly_agree    = "#1f77b4"
)[l_cat_stems]
cat_colours <- cat_colours[!is.na(cat_colours)]

ggplot(long_all,
       aes(x    = pct_directed,
           y    = interaction(time, .data[[QUESTION_COL]], sep = " – "),
           fill = category)) +
  geom_col(width = 0.65, colour = "white", linewidth = 0.3) +
  geom_vline(xintercept = 0, colour = "grey30", linewidth = 0.5) +
  scale_fill_manual(
    values = cat_colours,
    labels = str_to_title(str_replace_all(names(cat_colours), "_", " ")),
    guide  = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(
    labels = function(x) paste0(abs(round(x)), "%"),
    limits = c(-105, 105),
    breaks = seq(-100, 100, 25)
  ) +
  labs(
    title    = "Likert Questions: Response Distribution Pre vs. Post",
    subtitle = "Bars show % of respondents; left = disagree, right = agree",
    x        = NULL,
    y        = NULL,
    fill     = "Response"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 9)
  )
```

## Visualisation — Average Score Change

```{r likert-plot-avgscores, fig.height=max(4, nrow(results_likert) * 0.55 + 1.5)}
# Only render this plot if avg_score columns were detected
if (is.null(pre_score_col)) {
  message("No avg_score columns detected — skipping average score plot.")
} else {
ggplot(results_likert,
       aes(x = avg_diff,
           y = reorder(question, avg_diff),
           colour = significant)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "grey60") +
  geom_segment(aes(x = 0, xend = avg_diff,
                   y = question, yend = question),
               colour = "grey80", linewidth = 0.8) +
  geom_point(size = 4) +
  geom_text(aes(label = paste0(ifelse(avg_diff >= 0, "+", ""),
                               avg_diff)),
            hjust = ifelse(results_likert$avg_diff >= 0, -0.4, 1.4),
            size  = 3.2) +
  scale_colour_manual(
    values = c("TRUE" = "#e63946", "FALSE" = "#6c757d"),
    labels = c("TRUE" = "Significant (p<.05)", "FALSE" = "Not significant"),
    na.value = "grey80",
    name   = NULL
  ) +
  scale_x_continuous(
    labels = function(x) paste0(ifelse(x >= 0, "+", ""), x),
    expand = expansion(mult = 0.25)
  ) +
  labs(
    title    = "Likert Questions: Change in Average Score (Post − Pre)",
    subtitle = paste0("Scale: ", LIKERT_MIN, "–", LIKERT_MAX,
                      " | Positive = more favourable post-training"),
    x        = "Change in average score",
    y        = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position    = "top",
    panel.grid.major.y = element_blank()
  )
} # end if (!is.null(pre_score_col))
```

---

# Multi-Select Questions Analysis

> **Method:** Each response option is treated as a binary outcome (selected / not
> selected). A chi-square test compares the count of selections pre vs. post for
> each option within each question. Effect size is Cramér's V.

## Statistical Tests

```{r multiselect-stats}
pre_m_cols  <- paste0("pre_",  m_option_stems)
post_m_cols <- paste0("post_", m_option_stems)

# Detect total-respondents columns for converting counts to rates
pre_m_total_col  <- if (!is.na(m_total_stem)) paste0("pre_",  m_total_stem) else NULL
post_m_total_col <- if (!is.na(m_total_stem)) paste0("post_", m_total_stem) else NULL

run_multiselect_test <- function(row) {
  q_label <- if (QUESTION_COL %in% names(row) && !is.null(row[[QUESTION_COL]]))
               as.character(row[[QUESTION_COL]]) else paste0("Row_", rownames(row))
  pre_counts  <- suppressWarnings(as.numeric(row[pre_m_cols]))
  post_counts <- suppressWarnings(as.numeric(row[post_m_cols]))
  names(pre_counts)  <- m_option_stems
  names(post_counts) <- m_option_stems

  pre_total  <- if (!is.null(pre_m_total_col))
    suppressWarnings(as.numeric(row[[pre_m_total_col]])) else sum(pre_counts, na.rm=TRUE)
  post_total <- if (!is.null(post_m_total_col))
    suppressWarnings(as.numeric(row[[post_m_total_col]])) else sum(post_counts, na.rm=TRUE)

  # Run a chi-square on the full option distribution
  valid <- !is.na(pre_counts) & !is.na(post_counts)
  mat   <- rbind(pre_counts[valid], post_counts[valid])

  if (sum(valid) < 2 || any(rowSums(mat) == 0)) {
    return(data.frame(question=q_label,
                      option=NA, pre_pct=NA, post_pct=NA,
                      diff_pp=NA, chi_sq=NA, p_value=NA,
                      cramers_v=NA, significant=NA))
  }

  test <- tryCatch(chisq.test(mat, simulate.p.value = any(mat < 5)),
                   error=function(e) NULL)
  n    <- sum(mat)
  k    <- ncol(mat)
  cv   <- if (!is.null(test)) round(sqrt(test$statistic / (n*(k-1))), 3) else NA
  p    <- if (!is.null(test)) round(test$p.value, 4) else NA
  cs   <- if (!is.null(test)) round(test$statistic, 3) else NA

  # Per-option rows for detailed visualisation
  bind_rows(lapply(m_option_stems[valid], function(opt) {
    pc <- pre_counts[opt];  ptot <- if(is.na(pre_total) || pre_total==0) NA else pre_total
    qc <- post_counts[opt]; qtot <- if(is.na(post_total)||post_total==0) NA else post_total
    data.frame(
      question    = q_label,
      option      = opt,
      pre_pct     = round(pc / ptot * 100, 1),
      post_pct    = round(qc / qtot * 100, 1),
      diff_pp     = round((qc/qtot - pc/ptot)*100, 1),
      chi_sq      = cs,
      p_value     = p,
      cramers_v   = cv,
      significant = if (!is.na(p)) p < ALPHA else NA,
      stringsAsFactors = FALSE
    )
  }))
}

if (length(m_option_stems) == 0) {
  message("No option count columns detected in df_multiselect. Check column names.")
  results_multiselect <- data.frame()
} else {
  results_multiselect <- bind_rows(
    lapply(seq_len(nrow(df_multiselect)),
           function(i) run_multiselect_test(df_multiselect[i,]))
  )

  results_multiselect %>%
    mutate(
      pre_pct  = paste0(pre_pct, "%"),
      post_pct = paste0(post_pct, "%"),
      diff_pp  = paste0(ifelse(diff_pp >= 0, "+", ""), diff_pp, "pp"),
      sig_flag = case_when(is.na(significant) ~ "—",
                           significant        ~ "✓ Yes",
                           TRUE               ~ "No")
    ) %>%
    select(question, option, pre_pct, post_pct, diff_pp,
           chi_sq, p_value, cramers_v, sig_flag) %>%
    kable(
      col.names = c("Question","Option","Pre %","Post %","Change",
                    "χ²","p-value","Cramér's V","Significant?"),
      caption   = "Multi-Select: chi-square test on option selection rates"
    ) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"),
                  full_width=FALSE) %>%
    row_spec(which(results_multiselect$significant == TRUE),
             bold=TRUE, color="#e63946")
}
```

## Visualisation — Option Selection Rates

```{r multiselect-plot, fig.height=max(5, nrow(results_multiselect) * 0.45 + 2)}
if (nrow(results_multiselect) > 0) {
  plot_m <- results_multiselect %>%
    select(question, option, pre_pct, post_pct) %>%
    mutate(across(c(pre_pct, post_pct), as.numeric)) %>%
    pivot_longer(c(pre_pct, post_pct), names_to="time", values_to="pct") %>%
    mutate(time  = recode(time, pre_pct="Pre", post_pct="Post"),
           time  = factor(time, levels=c("Pre","Post")),
           label = paste0(question, " — ", str_to_title(str_replace_all(option,"_"," "))))

  sig_lbl_m <- results_multiselect %>%
    mutate(label_base = paste0(question," — ",
                                str_to_title(str_replace_all(option,"_"," "))),
           label      = ifelse(!is.na(significant) & significant,
                               paste0(label_base," *"), label_base)) %>%
    select(label_base, label)

  plot_m <- left_join(plot_m, sig_lbl_m, by=c("label"="label_base")) %>%
    mutate(label = coalesce(label.y, label.x)) %>%
    select(-label.x, -label.y)

  ggplot(plot_m, aes(x=pct, y=reorder(label, pct), fill=time)) +
    geom_col(position=position_dodge(width=0.7), width=0.6, colour="white") +
    geom_text(aes(label=paste0(round(pct,1),"%")),
              position=position_dodge(width=0.7), hjust=-0.15, size=3) +
    scale_fill_manual(values=PAL, name="Time point") +
    scale_x_continuous(labels=percent_format(scale=1),
                       expand=expansion(mult=c(0,0.2))) +
    labs(title="Multi-Select: % of Respondents Selecting Each Option",
         subtitle="* indicates statistically significant distribution shift (p < 0.05)",
         x="% Selected", y=NULL) +
    theme_minimal(base_size=11) +
    theme(legend.position="top", panel.grid.major.y=element_blank(),
          axis.text.y=element_text(size=9))
}
```

---

# Summary Table {#summary}

```{r summary-table}
# ── Combined results across all four question types ───────────────────────────
summary_s <- if (nrow(results_slider) > 0) {
  results_slider %>%
    transmute(
      question    = question,
      type        = "Perception Slider",
      metric      = paste0("mean: ", pre_mean, " → ", post_mean),
      change      = diff,
      p_value     = p_value,
      effect_size = paste0("d=", cohens_d),
      significant = significant
    )
} else data.frame()

summary_k <- results_knowledge %>%
  transmute(
    question    = question,
    type        = "Knowledge",
    metric      = paste0("% correct: ", pre_pct, "% → ", post_pct, "%"),
    change      = paste0(ifelse(pct_point_diff >= 0, "+", ""), pct_point_diff, "pp"),
    p_value     = p_value,
    effect_size = paste0(abs(pct_point_diff), "pp"),
    significant = significant
  )

summary_l <- results_likert %>%
  transmute(
    question    = question,
    type        = "Likert",
    metric      = if (!is.null(pre_score_col))
                    paste0("avg score: ", pre_avg, " → ", post_avg)
                  else paste0("χ²=", chi_sq),
    change      = if (!is.null(pre_score_col))
                    paste0(ifelse(avg_diff >= 0, "+", ""), avg_diff)
                  else NA_character_,
    p_value     = p_value,
    effect_size = paste0("V=", cramers_v),
    significant = significant
  )

# Multi-Select: one row per question (summarise across options)
summary_m <- if (nrow(results_multiselect) > 0) {
  results_multiselect %>%
    group_by(question) %>%
    summarise(
      n_sig   = sum(significant == TRUE, na.rm=TRUE),
      n_opts  = n(),
      p_min   = min(p_value, na.rm=TRUE),
      v_max   = max(cramers_v, na.rm=TRUE),
      .groups = "drop"
    ) %>%
    transmute(
      question    = question,
      type        = "Multi-Select",
      metric      = paste0(n_sig, "/", n_opts, " options changed"),
      change      = NA_character_,
      p_value     = round(p_min, 4),
      effect_size = paste0("V=", round(v_max, 3)),
      significant = p_min < ALPHA
    )
} else data.frame()

combined_summary <- bind_rows(summary_s, summary_k, summary_l, summary_m) %>%
  arrange(type, p_value)

combined_summary %>%
  mutate(
    sig_flag = case_when(
      significant == TRUE  ~ "✓",
      significant == FALSE ~ "—",
      TRUE                 ~ "?"
    )
  ) %>%
  select(type, question, metric, change, p_value, effect_size, sig_flag) %>%
  kable(
    col.names = c("Type", "Question", "Pre → Post", "Change",
                  "p-value", "Effect size", "Sig."),
    caption   = paste(REPORT_TITLE, "| Complete Results Summary"),
    align     = "llllccl"
  ) %>%
  kable_styling(bootstrap_options=c("striped","hover"), full_width=TRUE) %>%
  {
    tbl <- .
    type_order <- c("Perception Slider","Knowledge","Likert","Multi-Select")
    for (tp in type_order) {
      rows <- which(combined_summary$type == tp)
      if (length(rows) > 0)
        tbl <- pack_rows(tbl, tp, min(rows), max(rows), bold=TRUE)
    }
    tbl
  } %>%
  row_spec(which(combined_summary$significant == TRUE),
           bold=TRUE, color="#e63946")
```

---

# Session Info

```{r session-info, echo=FALSE}
# For reproducibility: records the R version and package versions used
sessionInfo()
```
